% ============================================================
% PART 3: Reproducibility, Learnings, Limitations, Future Work,
%         Conclusion, References
% ============================================================
% Merge after Part 2. Delete \end{document} from Part 1, then
% paste Part 2, then this file, then keep the \end{document} below.
% ============================================================

% ============================================================
% 8. REPRODUCIBILITY
% ============================================================
\section{Reproducibility and Experiment Tracking}

Reproducibility was a core design principle throughout this project.

\subsection{Fixed Experimental Constants}

To ensure valid single-variable comparisons, every experiment in the vR.P track shares:

\begin{table}[H]
\centering
\caption{Fixed Constants Across All vR.P Experiments}
\label{tab:constants}
\begin{tabular}{l l}
\toprule
\textbf{Constant} & \textbf{Value} \\
\midrule
Dataset & CASIA v2.0 (12,614 images) \\
Data split & 70/15/15 stratified \\
Random seed & 42 \\
Input resolution & 384$\times$384 \\
Base decoder & UNet (256/128/64/32/16) \\
Batch size & 16 \\
Optimizer & Adam ($\eta=10^{-3}$, wd=$10^{-5}$) \\
Framework & PyTorch 2.x + SMP \\
Hardware & Kaggle T4/P100 GPUs \\
Metric computation & Tampered images only \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Weights \& Biases Integration}

All experiments are logged to a centralized W\&B project~\cite{wandb2020} (\texttt{Tampered Image Detection \& Localization}), tracking:
\begin{itemize}[nosep]
    \item \textbf{Per-epoch:} \texttt{train\_loss}, \texttt{val\_loss}, \texttt{val\_pixel\_f1}, \texttt{val\_pixel\_iou}, \texttt{learning\_rate}
    \item \textbf{Final test:} \texttt{pixel\_f1}, \texttt{pixel\_iou}, \texttt{pixel\_auc}, \texttt{pixel\_precision}, \texttt{pixel\_recall}, \texttt{image\_accuracy}, \texttt{image\_macro\_f1}, \texttt{image\_roc\_auc}
    \item \textbf{Media:} \texttt{prediction\_examples} (W\&B Image panels)
\end{itemize}

The full interactive W\&B report is available at: \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}

\subsection{Infrastructure}

\begin{itemize}[nosep]
    \item \textbf{5 parallel Kaggle runner accounts} executing experiments via centralized configuration.
    \item \textbf{Papermill} for programmatic notebook execution from YAML experiment queues.
    \item \textbf{Centralized leaderboard notebook} that queries W\&B API and generates comparison tables, bar charts, and ablation progression plots automatically.
    \item \textbf{Google Colab T4 Optimised} notebooks for extended runs requiring more compute.
\end{itemize}

\subsection{Reproducibility Verification}

Two experiments were independently repeated to verify reproducibility:

\begin{table}[H]
\centering
\caption{Reproducibility Verification Results}
\label{tab:repro}
\begin{tabular}{l c c c}
\toprule
\textbf{Experiment} & \textbf{Run 1 F1} & \textbf{Run 2 F1} & \textbf{Match} \\
\midrule
vR.P.3 (ELA baseline) & 0.6920 & 0.6920 & Identical \\
vR.P.10 (CBAM) & 0.7277 & 0.7277 & Identical \\
\bottomrule
\end{tabular}
\end{table}

Both runs produced identical final metrics, confirming full reproducibility with seed 42 on Kaggle T4 GPUs.

% ============================================================
% 9. KEY LEARNINGS
% ============================================================
\section{Key Learnings}

The following lessons emerged from the full experimental journey across four research tracks and 60+ experiments:

\begin{enumerate}
    \item \textbf{Data leakage destroys credibility.} The Kaggle baseline (Track 2) appeared to work well ($\sim$89.75\% accuracy) but had inflated results due to test data leaking into training. Always audit data pipelines before trusting metrics.

    \item \textbf{Pretrained models are essential for small datasets.} Training from scratch on $\sim$10K images (Track 2) learns only low-level features and achieves Tampered F1 = 0.4101. A pretrained ResNet-34 on the same data achieves F1 = 0.4546, and with ELA reaches 0.6920.

    \item \textbf{Input representation matters more than everything else.} The ELA improvement (+23.74pp) exceeds all other improvements combined. Multi-quality RGB ELA adds another +10.45pp. No architectural change comes close.

    \item \textbf{Change one thing at a time.} Early attempts (Track 1) added multiple improvements simultaneously, leading to confusion and failure. Single-variable ablation is the only way to establish clear cause-and-effect relationships.

    \item \textbf{Negative results are valuable.} Knowing that DCT alone fails ($-$37.11pp), that TTA hurts localization ($-$5.32pp), and that Focal Loss is neutral (+0.03pp) prevents wasted effort in future work.

    \item \textbf{Verify the base approach supports all required tasks.} Track 3 (ETASR) achieved 90.23\% classification accuracy but was ultimately abandoned because it could not produce localization masks. This wasted significant effort.

    \item \textbf{Track experiments formally.} W\&B~\cite{wandb2020} made it possible to compare 60+ experiments systematically with reproducible metrics, automatic leaderboards, and visual comparisons. Without formal tracking, the project would have been chaotic.

    \item \textbf{The Flatten-Dense bottleneck is toxic.} In Track 3, the Flatten layer before Dense(256) created a 29.5M-parameter bottleneck that caused augmentation failure and training instability. Modern architectures avoid Flatten entirely using Global Average Pooling.

    \item \textbf{Freezing the encoder is the right default for small datasets.} Aggressive unfreezing (P.2, 23M trainable) overfit. Progressive unfreezing (P.8) confirmed that the best epoch was always in the frozen phase. Let the pretrained features do their job; only adapt BatchNorm.

    \item \textbf{Color information matters in forensic ELA.} Multi-Q RGB ELA (9ch, P.19, F1=0.7965) substantially outperformed Multi-Q grayscale ELA (3ch, P.15, F1=0.7329). The color channels encode which RGB channels were manipulated during forgery.
\end{enumerate}

% ============================================================
% 10. LIMITATIONS
% ============================================================
\section{Limitations}

\begin{enumerate}
    \item \textbf{Single dataset:} All results are on CASIA v2.0 only. Cross-dataset generalization to Columbia~\cite{hsu2006columbia}, Coverage~\cite{wen2016coverage}, CoMoFoD~\cite{tralic2013comofod}, and NIST~\cite{nist2019nimble} benchmarks is untested. Performance on out-of-distribution forgery types is unknown.

    \item \textbf{Fixed resolution:} 384$\times$384 input loses fine details, especially for small tampered regions. Higher resolution (512$\times$512 or 640$\times$640) may improve boundary precision but requires more GPU memory and longer training.

    \item \textbf{Binary segmentation only:} The model treats all tampering as one class. It does not distinguish between splicing and copy-move forgeries. Multi-class segmentation could provide more useful forensic information.

    \item \textbf{No robustness testing:} Behavior against post-processing attacks (JPEG recompression, Gaussian blur, noise addition, resizing) is not evaluated. Real-world tampered images often undergo multiple processing steps after manipulation.

    \item \textbf{No SRM noise features:} Spatial Rich Model~\cite{fridrich2012srm} noise extraction~--- a key technique in state-of-the-art systems like EMT-Net~\cite{emtnet2023} (AUC=0.987)~--- is not yet implemented.

    \item \textbf{ELA quality sensitivity:} The quality levels Q=75/85/95 were chosen by convention. A systematic quality sweep across all combinations was not performed.

    \item \textbf{Frozen encoder limitation:} Full fine-tuning with proper differential learning rates (lower LR for encoder, higher for decoder) and warmup scheduling is untested. This could potentially unlock better performance.

    \item \textbf{Diminishing returns ceiling:} After F1 $\sim$0.80, combining improvements shows diminishing or negative returns, suggesting the current architecture is approaching its ceiling on CASIA v2.0.
\end{enumerate}

% ============================================================
% 11. FUTURE WORK
% ============================================================
\section{Future Work}

\subsection{Immediate (High Confidence)}

\begin{itemize}[nosep]
    \item \textbf{Combine Multi-Q RGB ELA + CBAM with extended training:} The two best individual improvements have not been fully optimized together. The vR.P.30 series began this exploration, achieving F1=0.7762 vs.\ P.19's 0.7965, suggesting the combination needs different hyperparameters.
    \item \textbf{Differential learning rates:} Use lower LR ($10^{-5}$) for encoder BN layers and higher LR ($10^{-3}$) for decoder, with warmup scheduling.
    \item \textbf{Higher resolution input:} Test at 512$\times$512 and 640$\times$640 to improve boundary precision for small tampered regions.
\end{itemize}

\subsection{Medium-Term}

\begin{itemize}[nosep]
    \item \textbf{SRM noise filters:} Spatial Rich Model high-pass filters extract manipulation traces in the noise domain. Used by state-of-the-art systems.
    \item \textbf{YCbCr chrominance analysis:} Cb/Cr channels encode tamper traces better than luminance alone (reported 96.52\% on CASIA v2.0 with chrominance-only features).
    \item \textbf{JPEG quality augmentation:} Random re-compression during training to improve robustness against post-processing.
    \item \textbf{Cross-dataset evaluation:} Columbia~\cite{hsu2006columbia}, Coverage~\cite{wen2016coverage}, CoMoFoD~\cite{tralic2013comofod}, NIST~\cite{nist2019nimble} benchmarks to verify generalization.
\end{itemize}

\subsection{Long-Term}

\begin{itemize}[nosep]
    \item \textbf{Transformer-based encoders:} Swin Transformer or ConvNeXt may capture long-range manipulation patterns better than ResNet's local receptive fields.
    \item \textbf{Ensemble methods:} Combine predictions from Multi-Q ELA, CBAM, and augmentation variants using learned weights.
    \item \textbf{Multi-class segmentation:} Distinguish splicing from copy-move at the pixel level.
    \item \textbf{Robustness evaluation:} Systematic testing against JPEG recompression, blur, noise, and social media compression pipelines.
\end{itemize}

% ============================================================
% 12. CONCLUSION
% ============================================================
\section{Conclusion}

This report presented a systematic ablation study for image tamper detection and localization on the CASIA v2.0 dataset. Through 60+ controlled experiments across four research tracks, we demonstrated the following:

\begin{enumerate}
    \item \textbf{Input representation is the dominant factor.} Switching from raw RGB to Multi-Quality RGB ELA (9-channel) produced a +34.19 percentage point improvement in Pixel F1~--- far exceeding any architectural or training change.

    \item \textbf{The best model achieves Pixel F1 = 0.7965, IoU = 0.6615, and Pixel AUC = 0.9665}, using a UNet with a frozen pretrained ResNet-34 encoder and 9-channel multi-quality ELA input.

    \item \textbf{CBAM attention is the most parameter-efficient improvement}, adding +3.57pp Pixel F1 for only 11,402 parameters (0.36\% of trainable).

    \item \textbf{Single-variable ablation discipline} transforms ad-hoc experimentation into systematic scientific investigation, enabling clear cause-and-effect understanding.

    \item \textbf{Failures are as informative as successes.} Each failed research track (documentation-first, Kaggle reproduction, classification-only) contributed lessons that directly improved the final approach.
\end{enumerate}

The impact hierarchy established by this study:
\begin{equation}
\boxed{\text{Input representation} \gg \text{Attention} > \text{Training} > \text{Loss}}
\end{equation}
provides a clear roadmap for future work: invest in better input representations before trying more complex architectures.

All code, notebooks, and experiment logs are publicly available:
\begin{itemize}[nosep]
    \item GitHub: \url{https://github.com/The-Harsh-Vardhan/TIDAL-Tampered_Image_Detection_And_Localization}
    \item W\&B Report: \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}
\end{itemize}

% ============================================================
% REFERENCES
% ============================================================
\begin{thebibliography}{99}

\bibitem{dong2013casia}
J.~Dong, W.~Wang, and T.~Tan, ``CASIA Image Tampering Detection Evaluation Database,'' in \textit{Proc. IEEE China Summit and Int. Conf. on Signal and Information Processing}, pp.~422--426, 2013.

\bibitem{krawetz2007picture}
N.~Krawetz, ``A Picture's Worth\ldots Digital Image Analysis and Forensics,'' \textit{Hacker Factor Solutions}, 2007.

\bibitem{nagm2024etasr}
A.~Nagm \textit{et~al.}, ``Enhanced Image Tampering Detection using ELA and a CNN,'' \textit{Engineering, Technology \& Applied Science Research (ETASR)}, vol.~14, no.~1, pp.~12757--12762, 2024.

\bibitem{ronneberger2015unet}
O.~Ronneberger, P.~Fischer, and T.~Brox, ``U-Net: Convolutional Networks for Biomedical Image Segmentation,'' in \textit{Proc. MICCAI}, pp.~234--241, 2015.

\bibitem{he2016resnet}
K.~He, X.~Zhang, S.~Ren, and J.~Sun, ``Deep Residual Learning for Image Recognition,'' in \textit{Proc. IEEE CVPR}, pp.~770--778, 2016.

\bibitem{woo2018cbam}
S.~Woo, J.~Park, J.-Y.~Lee, and I.~S.~Kweon, ``CBAM: Convolutional Block Attention Module,'' in \textit{Proc. ECCV}, pp.~3--19, 2018.

\bibitem{yakubovskiy2019smp}
P.~Yakubovskiy, ``Segmentation Models PyTorch,'' GitHub, 2019. [Online]. Available: \url{https://github.com/qubvel/segmentation_models.pytorch}

\bibitem{bianchi2012jpeg}
T.~Bianchi and A.~Piva, ``Image Forgery Localization via Block-Grained Analysis of JPEG Artifacts,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~7, no.~3, pp.~1003--1017, 2012.

\bibitem{fridrich2012srm}
J.~Fridrich and J.~Kodovsky, ``Rich Models for Steganalysis of Digital Images,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~7, no.~3, pp.~868--882, 2012.

\bibitem{emtnet2023}
J.~Liu \textit{et~al.}, ``EMT-Net: Efficient Multi-Task Network for Image Manipulation Detection and Localization,'' in \textit{Proc. IEEE Int. Conf. on Multimedia and Expo (ICME)}, 2023.

\bibitem{deng2009imagenet}
J.~Deng, W.~Dong, R.~Socher, L.-J.~Li, K.~Li, and L.~Fei-Fei, ``ImageNet: A Large-Scale Hierarchical Image Database,'' in \textit{Proc. IEEE CVPR}, pp.~248--255, 2009.

\bibitem{milletari2016vnet}
F.~Milletari, N.~Navab, and S.-A.~Ahmadi, ``V-Net: Fully Convolutional Neural Networks for Volumetric Medical Image Segmentation,'' in \textit{Proc. Int. Conf. on 3D Vision (3DV)}, pp.~565--571, 2016.

\bibitem{lin2017focal}
T.-Y.~Lin, P.~Goyal, R.~Girshick, K.~He, and P.~Doll\'{a}r, ``Focal Loss for Dense Object Detection,'' in \textit{Proc. IEEE ICCV}, pp.~2980--2988, 2017.

\bibitem{kingma2015adam}
D.~P.~Kingma and J.~Ba, ``Adam: A Method for Stochastic Optimization,'' in \textit{Proc. ICLR}, 2015.

\bibitem{micikevicius2018mixed}
P.~Micikevicius \textit{et~al.}, ``Mixed Precision Training,'' in \textit{Proc. ICLR}, 2018.

\bibitem{zhou2018unetplusplus}
Z.~Zhou, M.~M.~Rahman~Siddiquee, N.~Tajbakhsh, and J.~Liang, ``UNet++: A Nested U-Net Architecture for Medical Image Segmentation,'' in \textit{Deep Learning in Medical Image Analysis}, pp.~3--11, 2018.

\bibitem{bayar2018constrained}
B.~Bayar and M.~C.~Stamm, ``Constrained Convolutional Neural Networks: A New Approach Towards General Purpose Image Manipulation Detection,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~13, no.~11, pp.~2691--2706, 2018.

\bibitem{wu2019mantranet}
Y.~Wu, W.~AbdAlmageed, and P.~Natarajan, ``ManTra-Net: Manipulation Tracing Network for Detection and Localization of Image Forgeries with Anomalous Features,'' in \textit{Proc. IEEE CVPR}, pp.~9543--9552, 2019.

\bibitem{wandb2020}
L.~Biewald, ``Experiment Tracking with Weights and Biases,'' 2020. [Online]. Available: \url{https://www.wandb.com}

\bibitem{hu2018senet}
J.~Hu, L.~Shen, and G.~Sun, ``Squeeze-and-Excitation Networks,'' in \textit{Proc. IEEE CVPR}, pp.~7132--7141, 2018.

\bibitem{hu2020span}
X.~Hu, Z.~Zhang, Z.~Jiang, S.~Chaudhuri, Z.~Yang, and R.~Nevatia, ``SPAN: Spatial Pyramid Attention Network for Image Manipulation Localization,'' in \textit{Proc. ECCV}, pp.~312--328, 2020.

\bibitem{wang2022objectformer}
J.~Wang, Z.~Wu, J.~Chen, X.~Han, A.~Shrivastava, S.~Lim, and Y.-G.~Jiang, ``ObjectFormer for Image Manipulation Detection and Localization,'' in \textit{Proc. IEEE CVPR}, pp.~2364--2373, 2022.

\bibitem{wu2022ifosn}
H.~Wu, J.~Zhou, J.~Tian, and J.~Liu, ``Robust Image Forgery Detection Against Transmission Over Online Social Networks,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~17, pp.~443--456, 2022.

\bibitem{hsu2006columbia}
Y.-F.~Hsu and S.-F.~Chang, ``Detecting Image Splicing Using Geometry Invariants and Camera Characteristics Consistency,'' in \textit{Proc. IEEE Int. Conf. on Multimedia and Expo}, pp.~549--552, 2006.

\bibitem{wen2016coverage}
B.~Wen, Y.~Zhu, R.~Subramanian, T.-T.~Ng, X.~Shen, and S.~Winkler, ``COVERAGE~--- A Novel Database for Copy-Move Forgery Detection,'' in \textit{Proc. IEEE Int. Conf. on Image Processing (ICIP)}, pp.~161--165, 2016.

\bibitem{tralic2013comofod}
D.~Tralic, I.~Zupancic, S.~Grgic, and M.~Grgic, ``CoMoFoD~--- New Database for Copy-Move Forgery Detection,'' in \textit{Proc. ELMAR}, pp.~49--54, 2013.

\bibitem{nist2019nimble}
NIST, ``Media Forensics Challenge (MFC) and Nimble Challenge,'' National Institute of Standards and Technology, 2019.

\bibitem{tan2019efficientnet}
M.~Tan and Q.~V.~Le, ``EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks,'' in \textit{Proc. ICML}, pp.~6105--6114, 2019.

\bibitem{loshchilov2017cosine}
I.~Loshchilov and F.~Hutter, ``SGDR: Stochastic Gradient Descent with Warm Restarts,'' in \textit{Proc. ICLR}, 2017.

\bibitem{ioffe2015batchnorm}
S.~Ioffe and C.~Szegedy, ``Batch Normalization: Accelerating Deep Network Training by Reducing Internal Covariate Shift,'' in \textit{Proc. ICML}, pp.~448--456, 2015.

\end{thebibliography}

\end{document}
