% ============================================================
% PART 2: Research Journey, Ablation Study, Results
% ============================================================
% This is a continuation. Merge after Part 1's \end{document} line.
% Delete the \end{document} from Part 1 and paste this content.
% ============================================================

% ============================================================
% 5. RESEARCH JOURNEY
% ============================================================
\section{Research Journey: Four Experimental Lineages}

This project did not follow a straight line. Each failed approach taught lessons that directly shaped the next attempt. We document all four tracks because the failures were as informative as the successes.

\subsection{Track 1: Documentation-First (v0x) --- Failed}

The project began with an extensive literature review. A survey of 21 research papers was conducted, and documentation was written before any code: architecture specifications, dataset analyses, training plans~--- all generated with AI assistance (Gemini Deep Research, ChatGPT, Codex). The philosophy was ``get the docs right, then build notebooks from them.''

This approach failed when too many ideas from research papers were added simultaneously, leading to untested complexity. There was little to no executable code until late versions.

\textbf{Lesson:} Documentation without experimentation leads to untested assumptions. Ideas must be validated incrementally.

\subsection{Track 2: Kaggle Notebook Reproduction (vK) --- Failed}

A Kaggle notebook (\texttt{image-detection-with-mask}) was discovered that appeared to solve the assignment task: a dual-head UNet with classification and segmentation outputs, reporting $\sim$89.75\% accuracy and Dice $\sim$0.57 on CASIA.

An audit revealed two fatal problems:
\begin{enumerate}[nosep]
    \item \textbf{Data leakage:} The test CSV pointed to the validation set. Reported ``test accuracy'' was actually validation accuracy.
    \item \textbf{Segmentation failure:} On a dataset where $\sim$40\% of images are authentic (all-zero masks), a model predicting all-zero masks scores Dice $\sim$0.58. The segmentation head was not learning meaningful features.
\end{enumerate}

After fixing the leakage, training from scratch on $\sim$10K RGB images with 31.6M parameters proved insufficient. Best honest Tampered F1 in this track: 0.4101 (v6.5).

\textbf{Lesson:} Always audit for data leakage. Training from scratch on small datasets with large models is insufficient~--- the model memorizes low-level features instead of learning forensic patterns.

\subsection{Track 3: ETASR Classification (vR) --- Archived}

The project pivoted to reproducing a published research paper~\cite{nagm2024etasr}: ``Enhanced Image Tampering Detection using ELA and a CNN'' which reported 96.21\% accuracy using a custom 2-layer CNN on $128 \times 128$ ELA images.

Seven honest ablation experiments (vR.1.1 through vR.1.7) were conducted:

\begin{table}[H]
\centering
\caption{ETASR Classification Track Results}
\label{tab:etasr}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l l c c c l}
\toprule
\textbf{Ver.} & \textbf{Change} & \textbf{Acc} & \textbf{F1} & \textbf{Params} & \textbf{Verdict} \\
\midrule
vR.1.1 & Honest baseline & 88.38 & .8805 & 29.5M & Baseline \\
vR.1.2 & Data augmentation & 85.53 & .8505 & 29.5M & Rejected \\
vR.1.3 & Class weights & 89.17 & .8889 & 29.5M & Positive \\
vR.1.4 & BatchNorm & 88.75 & .8852 & 29.5M & Neutral \\
vR.1.5 & LR scheduler & 88.96 & .8873 & 29.5M & Neutral \\
\textbf{vR.1.6} & \textbf{Deeper CNN} & \textbf{90.23} & \textbf{.9004} & \textbf{13.8M} & \textbf{Best} \\
vR.1.7 & Global Avg Pool & 89.17 & .8901 & 64K & Neutral \\
\bottomrule
\end{tabular}
\end{table}

Key findings from this track:
\begin{itemize}[nosep]
    \item \textbf{Paper gap:} 5.98pp below the claimed 96.21\%, never closed (likely due to dataset filtering: the paper used JPEG-only images).
    \item \textbf{The Flatten-Dense bottleneck is toxic:} The Flatten layer connecting the final conv map to Dense(256) created a 29.5M-parameter bottleneck (99\%+ of total parameters). It caused data augmentation failure (vR.1.2) because Flatten memorizes exact pixel positions~--- augmented images activate completely different neurons.
    \item \textbf{Architecture $>$ training tricks:} The deeper CNN (vR.1.6, +1.85pp) outperformed class weights + BatchNorm + LR scheduler combined (+0.58pp).
    \item \textbf{Critical limitation:} Classification-only. No pixel-level localization capability.
\end{itemize}

The track was archived because the assignment requires localization.

\textbf{Lesson:} Verify the base architecture supports all required output tasks before investing in ablation studies.

\subsection{Track 4: Pretrained Ablation Study (vR.P) --- Final}

Armed with lessons from all prior failures, the final lineage used:
\begin{itemize}[nosep]
    \item \textbf{Pretrained encoder} (ResNet-34 on ImageNet) to avoid the small-dataset problem.
    \item \textbf{UNet segmentation} for pixel-level localization.
    \item \textbf{ELA preprocessing} (from ETASR track insight) as forensic input.
    \item \textbf{Single-variable ablation:} exactly one change per experiment.
    \item \textbf{Weights \& Biases}~\cite{wandb2020} tracking for all experiments.
    \item \textbf{Reproducibility protocol:} seed 42, identical hardware, two verification runs.
\end{itemize}

% ============================================================
% 6. COMPLETE ABLATION STUDY
% ============================================================
\section{Complete Ablation Study}

The ablation study is the core contribution of this work. Each experiment changed \textit{exactly one variable} from a known baseline, enabling clear cause-and-effect analysis.

\subsection{Full Results Table}

Table~\ref{tab:full_results} presents the complete results from all experiments in the vR.P track.

\begin{table*}[t]
\centering
\caption{Complete Experiment Results from the vR.P Ablation Study. Each row represents a single-variable change. Bold entries indicate the best value in each column. Verdicts: Strong+ = major improvement; Positive = meaningful improvement; Neutral = within noise; Negative = degradation.}
\label{tab:full_results}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l l c c c c c c c l}
\toprule
\textbf{Version} & \textbf{Modification} & \textbf{Px F1} & \textbf{IoU} & \textbf{Px AUC} & \textbf{Img Acc} & \textbf{Img F1} & \textbf{Img AUC} & \textbf{Params} & \textbf{Verdict} \\
\midrule
P.0 & Dataset baseline (no proper GT) & 0.3749 & 0.2307 & 0.8486 & 70.63\% & 0.6814 & 0.7860 & 3.17M & Baseline \\
P.1 & Proper GT masks & 0.4546 & 0.2942 & 0.8509 & 70.15\% & 0.6867 & 0.7785 & 3.17M & Baseline \\
P.1.5 & AMP + TF32 speed opts & 0.4227 & 0.2680 & 0.8560 & 71.05\% & 0.7016 & 0.7980 & 3.17M & Neutral \\
P.2 & Aggressive encoder unfreeze & 0.5117 & 0.3439 & 0.8688 & 69.04\% & 0.6673 & 0.7196 & 23.0M & Positive \\
\textbf{P.3} & \textbf{ELA input (replacing RGB)} & \textbf{0.6920} & 0.5291 & 0.9528 & 86.79\% & 0.8560 & 0.9502 & 3.17M & \textbf{Strong+} \\
P.4 & 4-channel RGB+ELA fusion & 0.7053 & 0.5447 & 0.9433 & 84.42\% & 0.8322 & 0.9229 & 3.17M & Neutral \\
P.5 & ResNet-50 encoder (RGB) & 0.5137 & 0.3456 & 0.8828 & 72.00\% & 0.7143 & 0.8126 & 9.0M & Positive \\
P.6 & EfficientNet-B0 encoder (RGB) & 0.5217 & 0.3529 & 0.8708 & 70.68\% & 0.6950 & 0.7801 & 2.24M & Positive \\
P.7 & Extended training (50 epochs) & 0.7154 & 0.5569 & 0.9504 & 87.37\% & 0.8637 & 0.9433 & 3.17M & Positive \\
P.8 & Progressive encoder unfreeze & 0.6985 & 0.5367 & 0.9541 & 87.59\% & 0.8650 & 0.9578 & 3.17M & Neutral \\
P.9 & Focal + Dice loss & 0.6923 & 0.5294 & 0.9323 & 87.16\% & 0.8606 & 0.9076 & 3.17M & Neutral \\
\textbf{P.10} & \textbf{CBAM attention in decoder} & 0.7277 & 0.5719 & 0.9573 & 87.32\% & 0.8615 & \textbf{0.9633} & 3.18M & \textbf{Positive} \\
P.12 & Data augmentation & 0.6968 & 0.5347 & 0.9502 & \textbf{88.48\%} & \textbf{0.8756} & 0.9427 & 3.17M & Neutral \\
P.14b & Test-time augmentation & 0.6388 & 0.4693 & \textbf{0.9618} & 87.43\% & 0.8619 & 0.9610 & 3.17M & Negative \\
\textbf{P.15} & \textbf{Multi-Q ELA (gray, Q=75/85/95)} & 0.7329 & 0.5785 & 0.9608 & 87.53\% & 0.8660 & 0.9423 & 3.17M & \textbf{Positive} \\
P.16 & DCT spatial map (replacing ELA) & 0.3209 & 0.1911 & 0.7778 & 61.60\% & 0.5678 & 0.6204 & 3.17M & Negative \\
P.17 & ELA + DCT fusion (6-channel) & 0.7302 & 0.5751 & 0.9431 & 87.06\% & 0.8589 & 0.9462 & 3.17M & Positive \\
\textbf{P.19} & \textbf{Multi-Q RGB ELA (9-channel)} & \textbf{0.7965} & \textbf{0.6615} & 0.9665 & --- & --- & --- & 3.17M & \textbf{Best} \\
P.27 & JPEG compression augmentation & 0.7523 & 0.6029 & --- & --- & --- & --- & 3.17M & Positive \\
P.28 & Cosine annealing LR scheduler & 0.7601 & 0.6132 & --- & --- & --- & --- & 3.17M & Positive \\
P.30 & Multi-Q ELA + CBAM (25ep) & 0.7714 & 0.6214 & --- & --- & --- & --- & 3.18M & Positive \\
P.30.1 & Multi-Q ELA + CBAM (50ep) & 0.7762 & 0.6344 & --- & --- & --- & --- & 3.18M & 2nd Best \\
P.30.2 & Multi-Q ELA + CBAM + unfreeze & 0.7721 & 0.6260 & --- & --- & --- & --- & 3.18M & Positive \\
P.30.3 & Multi-Q ELA + CBAM + focal & 0.7745 & --- & --- & --- & --- & --- & 3.18M & Positive \\
P.30.4 & Multi-Q ELA + CBAM + aug & 0.7745 & --- & --- & --- & --- & --- & 3.18M & Positive \\
\bottomrule
\end{tabular}
\end{table*}

\FloatBarrier

\subsection{Per-Metric Champions}

No single experiment dominates all metrics~--- a consistent finding across the study.

\begin{table}[H]
\centering
\caption{Per-Metric Champions}
\label{tab:champions}
\begin{tabular}{l l c}
\toprule
\textbf{Metric} & \textbf{Best Version} & \textbf{Value} \\
\midrule
\textbf{Pixel F1} & P.19 (Multi-Q RGB ELA) & \textbf{0.7965} \\
\textbf{Pixel IoU} & P.19 (Multi-Q RGB ELA) & \textbf{0.6615} \\
Pixel AUC & P.14b (TTA) & 0.9618 \\
Pixel Precision & P.8 (Prog. unfreeze) & 0.8857 \\
Image Accuracy & P.12 (Augmentation) & 88.48\% \\
Image Macro F1 & P.12 (Augmentation) & 0.8756 \\
Image AUC & P.10 (CBAM) & 0.9633 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablation by Category}

\subsubsection{Category 1: Input Representation (Dominant Factor)}

This was the most impactful category across the entire study. Table~\ref{tab:input_ablation} shows the input representation experiments.

\begin{table}[H]
\centering
\caption{Input Representation Ablation}
\label{tab:input_ablation}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l c c c}
\toprule
\textbf{Input Type} & \textbf{Channels} & \textbf{Px F1} & \textbf{$\Delta$ from P.1} \\
\midrule
RGB (P.1) & 3 & 0.4546 & --- \\
ELA gray $\to$ RGB (P.3) & 3 & 0.6920 & +0.2374 \\
RGB + ELA (P.4) & 4 & 0.7053 & +0.2507 \\
Multi-Q ELA gray (P.15) & 3 & 0.7329 & +0.2783 \\
ELA + DCT (P.17) & 6 & 0.7302 & +0.2756 \\
\textbf{Multi-Q RGB ELA (P.19)} & \textbf{9} & \textbf{0.7965} & \textbf{+0.3419} \\
DCT only (P.16) & 3 & 0.3209 & $-$0.1337 \\
\bottomrule
\end{tabular}
\end{table}

Key observations:
\begin{itemize}[nosep]
    \item \textbf{RGB $\to$ ELA (P.3):} +23.74pp~--- the single largest improvement. False positive rate dropped from 22.6\% to 2.7\%.
    \item \textbf{Grayscale $\to$ RGB ELA (P.15 vs P.19):} Keeping RGB color information in each ELA map was worth +6.36pp over grayscale. The color channels carry additional forensic information about which color channels were manipulated.
    \item \textbf{DCT-only (P.16):} Catastrophic $-$37.11pp failure. Block-level frequency coefficients without spatial ELA context are insufficient for the ResNet encoder, which expects spatially structured input.
\end{itemize}

\subsubsection{Category 2: Encoder Architecture}

\begin{table}[H]
\centering
\caption{Encoder Architecture Comparison (all on RGB input)}
\label{tab:encoder_ablation}
\begin{tabular}{l c c c}
\toprule
\textbf{Encoder} & \textbf{Train. Params} & \textbf{Px F1} & \textbf{$\Delta$ from P.1} \\
\midrule
ResNet-34 (P.1) & 3.17M & 0.4546 & --- \\
ResNet-50 (P.5) & 9.0M & 0.5137 & +0.0591 \\
EfficientNet-B0 (P.6) & 2.24M & 0.5217 & +0.0671 \\
ResNet-34 + ELA (P.3) & 3.17M & \textbf{0.6920} & \textbf{+0.2374} \\
\bottomrule
\end{tabular}
\end{table}

The hierarchy is stark: \textbf{ELA on ResNet-34} (3.17M params, F1=0.6920) vastly outperforms ResNet-50 (9M params, F1=0.5137) and EfficientNet-B0 (2.24M params, F1=0.5217) on RGB. Input representation matters far more than encoder capacity.

\subsubsection{Category 3: Training Strategies}

\begin{table}[H]
\centering
\caption{Training Strategy Ablation (all from ELA baseline P.3)}
\label{tab:training_ablation}
\begin{tabular}{l c c}
\toprule
\textbf{Strategy} & \textbf{Px F1} & \textbf{$\Delta$ from P.3} \\
\midrule
P.3 Baseline (25 epochs) & 0.6920 & --- \\
P.7 Extended (50 epochs) & 0.7154 & +0.0234 \\
P.8 Progressive unfreeze & 0.6985 & +0.0065 \\
P.12 Data augmentation & 0.6968 & +0.0048 \\
P.1.5 AMP/TF32 & 0.4227 & (neutral) \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[nosep]
    \item \textbf{Extended training (P.7):} P.3's best epoch was 25/25 (the max). With 50-epoch budget, the optimum was found at epoch 36, yielding +2.34pp.
    \item \textbf{Progressive unfreeze (P.8):} Best epoch was in the frozen stage (epoch 23), before any unfreezing. Confirms frozen+BN is the sweet spot.
    \item \textbf{Data augmentation (P.12):} Only +0.48pp pixel F1 but achieved the best image accuracy (88.48\%) and image macro F1 (0.8756). Augmentation helps classification robustness more than pixel-level precision.
\end{itemize}

\subsubsection{Category 4: Loss Function}

Focal+Dice (P.9) was essentially neutral (+0.03pp) versus BCE+Dice. Pixel AUC actually regressed by $-$2.05pp because Focal Loss concentrates predictions near 0/1 extremes ($\gamma=2$ suppresses the gradient for well-classified pixels), destroying probability calibration.

\subsubsection{Category 5: Evaluation Strategy}

Test-time augmentation (P.14b) averaged predictions across 4 views (original, H-flip, V-flip, both flips). This achieved the best Pixel AUC (0.9618) and lowest FP rate (1.2\%) but \textit{destroyed} Pixel F1 ($-$5.32pp). Averaging smooths sharp mask boundaries, hurting localization precision.

\subsection{Impact Hierarchy}

\begin{table}[H]
\centering
\caption{Complete Impact Hierarchy (Delta from ELA baseline P.3)}
\label{tab:impact}
\begin{tabular}{c l l c}
\toprule
\textbf{\#} & \textbf{Category} & \textbf{Change} & \textbf{$\Delta$ F1 (pp)} \\
\midrule
1 & Input & Multi-Q RGB ELA 9ch & +10.45 \\
2 & Input & Multi-Q ELA gray & +4.09 \\
3 & Input & ELA+DCT fusion & +3.82 \\
4 & Architecture & CBAM attention & +3.57 \\
5 & Training & Extended training & +2.34 \\
6 & Input & RGB+ELA fusion & +1.33 \\
7 & Training & Progressive unfreeze & +0.65 \\
8 & Training & Augmentation & +0.48 \\
9 & Loss & Focal+Dice & +0.03 \\
10 & Evaluation & TTA & $-$5.32 \\
11 & Input & DCT-only & $-$37.11 \\
\bottomrule
\end{tabular}
\end{table}

The hierarchy is clear:
\begin{equation}
\boxed{\text{Input} \gg \text{Attention} > \text{Training} > \text{Loss}}
\end{equation}

\subsection{Parameter Efficiency Analysis}

CBAM provides the highest improvement density of any modification:
\begin{equation}
    \text{Efficiency}_{\text{CBAM}} = \frac{3.57\,\text{pp}}{11{,}402\,\text{params}} = 0.313\,\text{pp per 1K params}
\end{equation}

By contrast:
\begin{equation}
    \text{Efficiency}_{\text{ResNet-50}} = \frac{5.91\,\text{pp}}{5{,}830{,}000\,\text{extra params}} = 0.001\,\text{pp per 1K params}
\end{equation}

CBAM is $\sim$300$\times$ more parameter-efficient than upgrading the encoder.

\subsection{Diminishing Returns}

After Pixel F1 $\sim$0.78, combining improvements shows diminishing returns:
\begin{itemize}[nosep]
    \item P.19 alone: F1 = 0.7965
    \item P.30.1 (P.19 + CBAM + 50ep): F1 = 0.7762 ($<$ P.19)
    \item P.30.4 (P.19 + CBAM + aug): F1 = 0.7745 ($<$ P.19)
\end{itemize}

This suggests the model is approaching the ceiling of what the dataset and architecture can support. The 9-channel RGB ELA representation may already capture most of the forensic signal available in CASIA v2.0.

% ============================================================
% 7. ERROR ANALYSIS
% ============================================================
\section{Error Analysis}

\subsection{False Positive and False Negative Rates}

\begin{table}[H]
\centering
\caption{Image-Level Error Rates Across Configurations}
\label{tab:error_rates}
\begin{tabular}{l c c c}
\toprule
\textbf{Config} & \textbf{FP Rate} & \textbf{FN Rate} & \textbf{$\Delta$FP} \\
\midrule
RGB baseline (P.1) & 22.6\% & 40.4\% & --- \\
ELA baseline (P.3) & 2.7\% & 28.6\% & $-$19.9pp \\
CBAM (P.10) & 2.0\% & 28.3\% & $-$0.7pp \\
Augmentation (P.12) & 2.6\% & 24.6\% & $-$0.1pp \\
TTA (P.14b) & 1.2\% & 29.3\% & $-$1.5pp \\
\bottomrule
\end{tabular}
\end{table}

ELA reduces the false positive rate by 19.9 percentage points (22.6\% $\to$ 2.7\%). CBAM further drops it to 2.0\%. The remaining errors are primarily false negatives.

\subsection{Failure Modes}

The model struggles with three specific failure modes:
\begin{enumerate}[nosep]
    \item \textbf{Subtle copy-move forgeries:} When source and target textures are very similar, the ELA residuals are nearly identical, providing no forensic signal.
    \item \textbf{Very small tampered regions} ($<$2\% of image area): The 384$\times$384 resolution and the encoder's downsampling lose fine boundary details for tiny manipulations.
    \item \textbf{Multi-generation JPEG:} Images re-saved multiple times have uniform compression artifacts everywhere, destroying the differential ELA signal between authentic and tampered regions.
\end{enumerate}

\subsection{Precision-Recall Trade-off}

The best model (P.15 detailed metrics available) achieves:

\begin{table}[H]
\centering
\caption{Complete Metrics for Grayscale-ELA Model (P.15)}
\label{tab:detailed_metrics}
\begin{tabular}{l c l}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Interpretation} \\
\midrule
Pixel F1 & 0.7329 & Primary metric \\
Pixel IoU & 0.5785 & Stricter overlap \\
Pixel AUC & 0.9608 & Discrimination quality \\
Pixel Precision & 0.8409 & Few false alarms \\
Pixel Recall & 0.6496 & Misses some regions \\
Image Accuracy & 87.53\% & Detection accuracy \\
Image Macro F1 & 0.8660 & Balanced class perf \\
Image ROC-AUC & 0.9423 & Ranking quality \\
\bottomrule
\end{tabular}
\end{table}

The model is conservative: high precision (0.8409) means it rarely flags authentic regions as tampered, but moderate recall (0.6496) means it misses about 35\% of tampered pixels. This trade-off is acceptable for forensic applications where false accusations are costly.

% ============================================================
% EXPERIMENT LINEAGE DIAGRAM
% ============================================================
\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    node distance=0.55cm and 1.0cm,
    basenode/.style={draw, rounded corners=3pt, minimum height=0.7cm, minimum width=2.5cm, align=center, font=\scriptsize, thick},
    bestnode/.style={basenode, fill=green!25, line width=1.5pt},
    posnode/.style={basenode, fill=blue!10},
    negnode/.style={basenode, fill=red!15},
    neutnode/.style={basenode, fill=gray!10},
    mainflow/.style={-{Stealth[length=2.5mm]}, thick, black},
    branchflow/.style={-{Stealth[length=2mm]}, thick, black!60},
    delta/.style={font=\tiny, fill=white, inner sep=1pt, text=black!70},
]
% Main trunk
\node[neutnode] (p0) {P.0 RGB Baseline\\\textbf{F1 = 0.3749}};
\node[neutnode, below=0.5cm of p0] (p1) {P.1 Dataset Fix\\\textbf{F1 = 0.4546}};
\node[posnode, below=0.5cm of p1] (p3) {P.3 ELA Input\\\textbf{F1 = 0.6920}};
\node[posnode, below=0.5cm of p3] (p15) {P.15 Multi-Q ELA (gray)\\\textbf{F1 = 0.7329}};
\node[bestnode, below=0.5cm of p15] (p19) {P.19 Multi-Q RGB ELA 9ch\\\textbf{F1 = 0.7965}};
% Deltas on main trunk
\draw[mainflow] (p0) -- node[delta, right=2pt] {+0.0797} (p1);
\draw[mainflow] (p1) -- node[delta, right=2pt] {\textcolor{green!50!black}{+0.2374}} (p3);
\draw[mainflow] (p3) -- node[delta, right=2pt] {+0.0409} (p15);
\draw[mainflow] (p15) -- node[delta, right=2pt] {+0.0636} (p19);
% Branches from P.3
\node[posnode, right=1.8cm of p3] (p10) {P.10 CBAM Attention\\\textbf{F1 = 0.7277}};
\node[posnode, above=0.5cm of p10] (p7) {P.7 Extended 50ep\\\textbf{F1 = 0.7154}};
\node[negnode, below=0.5cm of p10] (p16) {P.16 DCT Only\\\textbf{F1 = 0.3209}};
\draw[branchflow] (p3.east) -- node[delta, above=1pt] {+0.0234} (p7.west);
\draw[branchflow] (p3.east) -- node[delta, above=1pt] {+0.0357} (p10.west);
\draw[branchflow] (p3.east) -- node[delta, below=1pt] {\textcolor{red!70!black}{$-$0.3711}} (p16.west);
% Branches from P.19
\node[posnode, right=1.8cm of p19] (p30) {P.30 + CBAM\\\textbf{F1 = 0.7714}};
\node[posnode, right=1.0cm of p30] (p301) {P.30.1 + 50ep\\\textbf{F1 = 0.7762}};
\draw[branchflow] (p19.east) -- (p30.west);
\draw[branchflow] (p30.east) -- (p301.west);
% Labels
\node[font=\scriptsize\bfseries, text=green!50!black, left=0.15cm of p19] {BEST};
% Legend
\node[anchor=north west, font=\scriptsize\bfseries] at ([xshift=4.5cm, yshift=0.2cm]p0.north east) (lt) {Legend:};
\node[bestnode, minimum width=1.1cm, minimum height=0.3cm, right=0.15cm of lt, font=\tiny] {Best};
\node[posnode, minimum width=1.1cm, minimum height=0.3cm, right=2.3cm of lt, font=\tiny] {Positive};
\node[negnode, minimum width=1.1cm, minimum height=0.3cm, right=4.3cm of lt, font=\tiny] {Negative};
\end{tikzpicture}
\caption{Experiment lineage diagram for the vR.P research track, showing evolution from RGB baseline (P.0) to best model (P.19). Delta annotations on arrows show Pixel F1 change. The largest single improvement (+0.2374) was switching from RGB to ELA.}
\label{fig:lineage}
\end{figure*}

\FloatBarrier
