\documentclass[11pt, twocolumn]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{float}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{placeins}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[all]{hypcap}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit, calc, decorations.pathreplacing, matrix}

\hypersetup{colorlinks=true, linkcolor=blue!70!black, citecolor=green!50!black, urlcolor=blue!60!black}

\title{\LARGE \textbf{Image Tampering Detection and Localization:\\A Systematic Ablation Study Using Deep Learning} \vspace{0.3cm} \\ \large Detailed Technical Report}

\author{
Harsh Vardhan \\
Big Vision Internship Assignment \\
\texttt{GitHub:} \url{https://github.com/The-Harsh-Vardhan/TIDAL-Tampered_Image_Detection_And_Localization} \\
\texttt{W\&B:} \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}
}

\date{}

\begin{document}
\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}
Digital images can be manipulated in ways invisible to the human eye. Detecting such tampering and pinpointing exactly which pixels were altered is a critical challenge in image forensics. This report presents a systematic ablation study for pixel-level image tamper detection and localization on the CASIA~v2.0 benchmark dataset (12,614 images: 7,491 authentic + 5,123 tampered). The project explored four experimental lineages over 60+ controlled experiments, evolving from documentation-first planning through Kaggle notebook reproduction, research paper baseline recreation, and finally a pretrained UNet segmentation pipeline.

The final system uses a UNet encoder-decoder with a frozen ResNet-34 encoder (ImageNet pretrained), multi-quality Error Level Analysis (ELA) preprocessing, and CBAM attention modules. Through strict single-variable ablation (one change per experiment), we established that \textbf{input representation dominates all other factors}~--- switching from RGB to multi-quality RGB ELA (9-channel) was worth +34.19 percentage points (pp) in Pixel~F1, more than all other improvements combined. The best model achieves \textbf{Pixel~F1 = 0.7965}, \textbf{IoU = 0.6615}, and \textbf{Pixel~AUC = 0.9665}. The impact hierarchy is clear: \textbf{Input representation $>$ Attention mechanisms $>$ Training configuration $>$ Loss function}.
\end{abstract}

\textbf{Keywords:} Image tampering detection, image forgery localization, error level analysis, UNet segmentation, ablation study, CASIA dataset, CBAM attention.

% ============================================================
% 1. INTRODUCTION
% ============================================================
\section{Introduction}

Digital images are ubiquitous in journalism, law enforcement, scientific research, and social media. However, modern image editing tools such as Adobe Photoshop, GIMP, and AI-based inpainting make it easy to manipulate photographs in ways that are nearly undetectable by the human eye. Common types of tampering include:

\begin{itemize}[nosep]
    \item \textbf{Splicing:} Copying a region from one image and pasting it into another.
    \item \textbf{Copy-move:} Duplicating a part of an image within the same image, often to hide or fabricate content.
    \item \textbf{Inpainting:} Using AI algorithms to fill in or replace regions of an image.
\end{itemize}

The task of \textit{image forgery detection} (classifying an image as authentic or tampered) is useful, but \textit{image forgery localization} (producing a pixel-level mask of tampered regions) is far more valuable for forensic analysis. This project tackles both tasks simultaneously: given an input image, the system produces a binary mask highlighting tampered pixels, from which image-level classification is derived.

\subsection{Problem Formulation}

Given an input image $\mathbf{X} \in \mathbb{R}^{H \times W \times 3}$, the goal is to learn a mapping function $f_\theta$ that produces a pixel-level probability map:
\begin{equation}
    \hat{\mathbf{M}} = f_\theta(\mathbf{X}) \in [0, 1]^{H \times W}
\end{equation}
where $\hat{\mathbf{M}}_{i,j}$ represents the probability that pixel $(i,j)$ has been tampered with. The ground truth mask $\mathbf{M} \in \{0, 1\}^{H \times W}$ is a binary map where 1 indicates tampered pixels and 0 indicates authentic pixels.

Image-level classification is derived from the segmentation output:
\begin{equation}
    \hat{y} = \begin{cases} 1 & \text{if } \max_{i,j} \hat{\mathbf{M}}_{i,j} > \tau \\ 0 & \text{otherwise} \end{cases}
\end{equation}
where $\tau = 0.5$ is the binarization threshold.

\subsection{Challenges}

Building an effective tamper localization system faces several fundamental difficulties:

\begin{enumerate}[nosep]
    \item \textbf{Extreme pixel-level class imbalance:} Tampered regions often occupy less than 2\% of all pixels in an image. A model that predicts all pixels as authentic achieves $>$98\% pixel accuracy while being completely useless.
    \item \textbf{Forgery sophistication:} Skilled forgers adjust lighting, color balance, and compression to make edits blend seamlessly with surrounding content.
    \item \textbf{JPEG compression artifacts:} Re-saving images as JPEG introduces compression artifacts that can either mask tampering traces or create false forensic signals.
    \item \textbf{Small dataset size:} Forensic datasets like CASIA v2.0 contain only $\sim$12K images~--- orders of magnitude smaller than ImageNet's 1.2M images~\cite{deng2009imagenet}, making training from scratch impractical.
\end{enumerate}

\subsection{Contributions}

The key contributions of this work are:
\begin{enumerate}[nosep]
    \item A complete pipeline from raw images to pixel-level tamper localization using multi-quality ELA preprocessing and pretrained UNet segmentation.
    \item A rigorous ablation study across 60+ experiments establishing that \textbf{input representation dominates all other factors}.
    \item Transparent documentation of four research tracks, including three failed approaches whose lessons directly shaped the final system.
    \item Full reproducibility via Weights~\&~Biases~\cite{wandb2020} experiment tracking, fixed random seeds, and open-source code.
\end{enumerate}

% ============================================================
% 2. RELATED WORK
% ============================================================
\section{Related Work}

\subsection{Traditional Forensic Methods}

Early forgery detection relied on hand-crafted features. Error Level Analysis (ELA)~\cite{krawetz2007picture} re-saves an image at a given JPEG quality and measures the pixel-wise difference from the original; tampered regions show inconsistent compression residuals. JPEG block artifact analysis~\cite{bianchi2012jpeg} exploits the 8$\times$8 block structure of JPEG compression to detect re-compressed regions. Noise-based methods analyze local noise variance~--- tampered regions composited from different sources often have mismatched noise patterns. Color Filter Array (CFA) analysis detects inconsistencies in the Bayer pattern interpolation. While interpretable, these methods struggle with sophisticated forgeries and produce high false-positive rates when used in isolation.

\subsection{CNN-Based Classification}

Deep learning transformed forgery detection by learning features directly from data. Bayar and Stamm~\cite{bayar2018constrained} proposed constrained convolutional layers that suppress image content and learn manipulation-specific features. Nagm et al.~\cite{nagm2024etasr} combined ELA preprocessing with a custom CNN for binary classification (tampered vs.\ authentic), reporting 96.21\% accuracy on CASIA v2.0 (JPEG-only subset). However, classification-only models cannot localize tampered regions~--- they answer ``is it tampered?'' but not ``where?''

\subsection{Segmentation-Based Localization}

To achieve pixel-level localization, researchers adopted encoder-decoder architectures. The UNet~\cite{ronneberger2015unet}, originally designed for biomedical image segmentation, proved effective for tamper localization through its skip connections that preserve fine spatial details. UNet++~\cite{zhou2018unetplusplus} improved boundary delineation through nested dense skip connections. Using pretrained encoders (e.g., ResNet~\cite{he2016resnet} trained on ImageNet~\cite{deng2009imagenet}) significantly improves performance on small forensic datasets by providing rich, generalizable feature representations.

ManTra-Net~\cite{wu2019mantranet} introduced a self-supervised anomaly detection approach for forgery localization that generalizes across manipulation types without per-dataset fine-tuning. SPAN~\cite{hu2020span} used spatial pyramid attention to capture multi-scale manipulation traces for more precise boundary localization.

\subsection{Attention Mechanisms}

Attention modules help models focus on the most relevant features. SE-Net~\cite{hu2018senet} introduced channel attention through squeeze-and-excitation blocks. CBAM~\cite{woo2018cbam} extended this with both channel-wise and spatial attention, teaching the model which feature maps and spatial locations are most important. In our experiments, CBAM provided one of the most parameter-efficient improvements (+3.57pp Pixel F1 for only 11K parameters).

\subsection{Multi-Stream and Frequency-Domain Methods}

State-of-the-art systems combine multiple input streams. EMT-Net~\cite{emtnet2023} uses Swin Transformers with RGB, noise maps, and frequency features, achieving AUC=0.987 on NIST. SRM (Spatial Rich Model)~\cite{fridrich2012srm} high-pass filters extract manipulation traces in the noise domain. ObjectFormer~\cite{wang2022objectformer} uses transformer-based architectures that learn object-level relationships between image patches to detect manipulated regions. IF-OSN~\cite{wu2022ifosn} combines frequency-domain analysis with online social network compression patterns for robust detection even after social media re-compression.

These multi-stream methods represent the current frontier but require significantly more computational resources and training data. Our work demonstrates that substantial performance can be achieved with simpler architectures when the input representation is well-designed.

% ============================================================
% 3. DATASET
% ============================================================
\section{Dataset}

\subsection{CASIA v2.0}

All experiments use the \textbf{CASIA v2.0} dataset~\cite{dong2013casia}, a widely-used benchmark for image forgery detection and localization.

\begin{table}[H]
\centering
\caption{CASIA v2.0 Dataset Summary}
\label{tab:dataset}
\begin{tabular}{l r}
\toprule
\textbf{Property} & \textbf{Value} \\
\midrule
Total images & 12,614 \\
Authentic images & 7,491 (59.4\%) \\
Tampered images & 5,123 (40.6\%) \\
\quad Copy-move forgeries & 3,295 (26.1\%) \\
\quad Splicing forgeries & 1,828 (14.5\%) \\
Ground truth & Binary masks \\
Image formats & JPEG, BMP, TIFF \\
Resolution range & Variable (resized to 384$\times$384) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data Splitting}

The dataset is split using a two-step stratified process with seed 42:

\begin{enumerate}[nosep]
    \item 70/30 split of the full dataset (stratified by authentic/tampered label).
    \item 50/50 split of the 30\% holdout into validation and test sets.
\end{enumerate}

\begin{table}[H]
\centering
\caption{Data Split Configuration}
\label{tab:split}
\begin{tabular}{l r r l}
\toprule
\textbf{Split} & \textbf{Count} & \textbf{Ratio} & \textbf{Purpose} \\
\midrule
Training & 8,829 & 70\% & Model optimization \\
Validation & 1,892 & 15\% & Early stopping, LR scheduling \\
Test & 1,893 & 15\% & Final reported metrics \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Preprocessing}

\begin{itemize}[nosep]
    \item All images resized to $384 \times 384$ pixels using bilinear interpolation.
    \item Ground truth masks resized with nearest-neighbor interpolation, then binarized at threshold 0.5.
    \item Per-channel normalization statistics computed from 500 randomly sampled training images.
\end{itemize}

\subsection{Data Leakage Discovery}

An important finding during the project was that the Kaggle variant of the CASIA dataset mixed ground truth mask images into the input data folder. A CNN trivially distinguished mask images (simple binary patterns) from real photographs, producing 100\% ``test accuracy.'' This result was discarded, and the discovery motivated strict pipeline audits in all subsequent experiments.

% ============================================================
% 4. METHODOLOGY
% ============================================================
\section{Methodology}

The final system consists of three components: (1) multi-quality ELA preprocessing, (2) a UNet segmentation model with a pretrained frozen encoder and CBAM attention, and (3) a training pipeline with carefully chosen loss functions and hyperparameters.

% ---- Architecture Diagram (vertical, single-column) ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.6cm,
    box/.style={draw, rounded corners, minimum height=0.75cm, minimum width=5.6cm, align=center, font=\footnotesize},
    outbox/.style={draw, rounded corners, minimum height=0.7cm, minimum width=2.4cm, align=center, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
]
\node[box, fill=blue!10] (input) {Input Image (384$\times$384$\times$3 RGB)};
\node[box, fill=orange!15, below=of input] (ela) {Multi-Quality ELA Preprocessing\\Q=75, Q=85, Q=95 $\times$ RGB = 9 channels};
\node[box, fill=green!15, below=of ela] (encoder) {ResNet-34 Encoder\\(ImageNet Pretrained, Frozen + BN Unfrozen)};
\node[box, fill=purple!12, below=of encoder] (decoder) {UNet Decoder (5 blocks)\\+ CBAM Attention + Skip Connections};
\node[box, fill=yellow!12, below=of decoder] (sigmoid) {Sigmoid Output (384$\times$384$\times$1)};
\node[outbox, fill=cyan!15, below left=0.55cm and -0.3cm of sigmoid] (mask) {Segmentation\\Mask};
\node[outbox, fill=cyan!15, below right=0.55cm and -0.3cm of sigmoid] (label_out) {Image Label\\(Auth / Tampered)};
\draw[arrow] (input) -- (ela);
\draw[arrow] (ela) -- (encoder);
\draw[arrow] (encoder) -- (decoder);
\draw[arrow] (decoder) -- (sigmoid);
\draw[arrow] (sigmoid.south) -- ++(-1.3,-0.25) -- (mask.north);
\draw[arrow] (sigmoid.south) -- ++(1.3,-0.25) -- (label_out.north);
\draw[arrow, dashed, gray] (encoder.east) -- ++(0.3,0) |- node[pos=0.25, right, font=\scriptsize, gray, align=left] {skip\\connections} (decoder.east);
\node[font=\scriptsize, gray, right=0.05cm of label_out] {max pixel prob.};
\end{tikzpicture}
\caption{Overall architecture of the tamper detection and localization pipeline.}
\label{fig:architecture}
\end{figure}

\subsection{Error Level Analysis (ELA)}

ELA is a forensic preprocessing technique that reveals JPEG compression inconsistencies. The core idea is that authentic regions of an image compress uniformly, while tampered regions~--- composited from different sources or saved at different quality levels~--- show inconsistent compression residuals.

\subsubsection{Mathematical Formulation}

Given an input image $\mathbf{X} \in \mathbb{R}^{H \times W \times 3}$ and a JPEG quality level $Q \in \{1, \ldots, 100\}$, the ELA map is computed as:
\begin{equation}
    \text{ELA}_Q(\mathbf{X}) = \left| \mathbf{X} - \text{JPEG}_Q(\mathbf{X}) \right| \cdot s
    \label{eq:ela}
\end{equation}
where $\text{JPEG}_Q(\cdot)$ denotes JPEG re-compression at quality $Q$, and $s$ is a brightness scaling factor such that the maximum channel difference maps to 255:
\begin{equation}
    s = \frac{255}{\max_{i,j,c} \left| X_{i,j,c} - \text{JPEG}_Q(\mathbf{X})_{i,j,c} \right| + \epsilon}
\end{equation}
with $\epsilon = 10^{-8}$ for numerical stability.

\subsubsection{Multi-Quality RGB ELA (Best Variant)}

Instead of a single quality level, the best configuration uses three quality levels, each producing an RGB ELA map (3 channels). These are stacked to form a 9-channel input:
\begin{equation}
    \mathbf{X}_{\text{MQ-ELA}} = \left[ \text{ELA}_{75}(\mathbf{X}) \,\|\, \text{ELA}_{85}(\mathbf{X}) \,\|\, \text{ELA}_{95}(\mathbf{X}) \right] \in \mathbb{R}^{H \times W \times 9}
\end{equation}
where $\|$ denotes channel-wise concatenation.

Each quality level captures different forensic information:

\begin{table}[H]
\centering
\caption{Multi-Quality ELA Channel Statistics (from P.15 training set)}
\label{tab:ela_stats}
\begin{tabular}{l c c l}
\toprule
\textbf{Quality} & \textbf{Mean} & \textbf{Std} & \textbf{Sensitivity} \\
\midrule
Q=75 (aggressive) & 0.0684 & 0.0656 & Strong manipulations \\
Q=85 (balanced) & 0.0605 & 0.0604 & Medium edits \\
Q=95 (gentle) & 0.0402 & 0.0471 & Subtle changes \\
\bottomrule
\end{tabular}
\end{table}

The three channels carry meaningfully different statistics~--- lower $Q$ produces larger residuals, confirming they provide distinct, non-redundant forensic information.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.45cm,
    box/.style={draw, rounded corners, minimum height=0.65cm, minimum width=2.6cm, align=center, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
]
\node[box, fill=orange!10] (q75) {Re-save at Q=75};
\node[box, fill=orange!10, right=0.25cm of q75] (q85) {Re-save at Q=85};
\node[box, fill=orange!10, right=0.25cm of q85] (q95) {Re-save at Q=95};
\node[box, fill=blue!10, above=0.5cm of q85] (orig) {Original Image (RGB)};
\node[box, fill=yellow!15, below=0.5cm of q85] (diff) {$|\text{Original} - \text{Recompressed}|$};
\node[box, fill=green!15, below=0.4cm of diff] (stack) {Stack $\rightarrow$ 9-Channel Tensor};
\node[box, fill=purple!10, below=0.4cm of stack] (model) {Feed to UNet Model};
\draw[arrow] (orig) -- (q75);
\draw[arrow] (orig) -- (q85);
\draw[arrow] (orig) -- (q95);
\draw[arrow] (q75) -- (diff);
\draw[arrow] (q85) -- (diff);
\draw[arrow] (q95) -- (diff);
\draw[arrow] (diff) -- (stack);
\draw[arrow] (stack) -- (model);
\end{tikzpicture}
\caption{Multi-Quality RGB ELA preprocessing pipeline.}
\label{fig:ela_pipeline}
\end{figure}

\subsubsection{ELA Algorithm}

\begin{algorithm}[H]
\caption{Multi-Quality RGB ELA Computation}
\label{alg:ela}
\begin{algorithmic}[1]
\Require Input image $\mathbf{X}$, quality levels $\mathcal{Q} = \{75, 85, 95\}$
\Ensure 9-channel ELA tensor $\mathbf{X}_{\text{MQ-ELA}}$
\State $\text{channels} \gets [\,]$
\For{$Q \in \mathcal{Q}$}
    \State $\mathbf{X}_Q \gets \text{JPEG\_recompress}(\mathbf{X}, Q)$
    \State $\mathbf{D}_Q \gets |\mathbf{X} - \mathbf{X}_Q|$
    \State $s_Q \gets 255 \,/\, (\max(\mathbf{D}_Q) + \epsilon)$
    \State $\text{channels.append}(\mathbf{D}_Q \cdot s_Q)$ \Comment{3 RGB channels}
\EndFor
\State $\mathbf{X}_{\text{MQ-ELA}} \gets \text{concatenate}(\text{channels}, \text{axis}=\text{channel})$
\State \Return $\mathbf{X}_{\text{MQ-ELA}}$
\end{algorithmic}
\end{algorithm}

\subsection{Model Architecture}

\subsubsection{UNet with ResNet-34 Encoder}

The model uses the Segmentation Models PyTorch (SMP) library's~\cite{yakubovskiy2019smp} UNet implementation with a ResNet-34~\cite{he2016resnet} encoder.

The ResNet-34 encoder uses residual connections:
\begin{equation}
    \mathbf{h}_l = \sigma(\mathbf{W}_l \ast \mathbf{h}_{l-1} + \mathbf{h}_{l-1})
\end{equation}
where $\ast$ denotes convolution, $\sigma$ is the activation function (ReLU), and the identity shortcut $\mathbf{h}_{l-1}$ allows gradients to flow directly through the network.

\textbf{Freeze strategy:} All encoder convolutional weights are frozen. Only BatchNorm layers remain unfrozen (trainable scale $\gamma$ and shift $\beta$ parameters). This allows BN statistics to adapt from ImageNet's natural image distribution to the ELA forensic domain:
\begin{equation}
    \hat{x} = \gamma \cdot \frac{x - \mu_{\text{batch}}}{\sqrt{\sigma^2_{\text{batch}} + \epsilon}} + \beta
\end{equation}

The first convolutional layer is modified to accept 9 input channels (from 3) by repeating the pretrained weights 3 times.

\begin{table}[H]
\centering
\caption{UNet Decoder Configuration}
\label{tab:decoder}
\begin{tabular}{c c c c}
\toprule
\textbf{Block} & \textbf{Channels} & \textbf{Input Scale} & \textbf{Skip From} \\
\midrule
1 & 256 & 1/16 & Encoder stage 4 \\
2 & 128 & 1/8 & Encoder stage 3 \\
3 & 64 & 1/4 & Encoder stage 2 \\
4 & 32 & 1/2 & Encoder stage 1 \\
5 & 16 & 1/1 & Input \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Model Parameter Summary}
\label{tab:params}
\begin{tabular}{l r r}
\toprule
\textbf{Component} & \textbf{Parameters} & \textbf{Trainable} \\
\midrule
Encoder (ResNet-34) & 21.3M & BN only \\
Decoder (5 blocks) & $\sim$3.17M & Yes \\
CBAM modules & 11,402 & Yes \\
Segmentation head & $\sim$1K & Yes \\
\midrule
\textbf{Total} & \textbf{24.4M} & \textbf{3.17M (13\%)} \\
\bottomrule
\end{tabular}
\end{table}

This freeze strategy was validated empirically: P.2's aggressive unfreeze (23M trainable) overfit with a 1:2,615 data-to-parameter ratio. P.8's progressive unfreeze confirmed the best epoch occurred before any unfreezing stage.

\subsubsection{CBAM Attention Module}

CBAM~\cite{woo2018cbam} is injected into all 5 UNet decoder blocks. It applies two sequential attention operations:

\textbf{Channel Attention.} Given a feature map $\mathbf{F} \in \mathbb{R}^{C \times H \times W}$:
\begin{equation}
    \mathbf{M}_c(\mathbf{F}) = \sigma\!\Big(\text{MLP}\!\big(\text{AvgPool}(\mathbf{F})\big) + \text{MLP}\!\big(\text{MaxPool}(\mathbf{F})\big)\Big)
\end{equation}
where $\sigma$ is the sigmoid function and MLP is a shared two-layer network with bottleneck ratio $r=16$:
\begin{equation}
    \text{MLP}(\mathbf{z}) = \mathbf{W}_1 \cdot \text{ReLU}(\mathbf{W}_0 \cdot \mathbf{z})
\end{equation}
with $\mathbf{W}_0 \in \mathbb{R}^{C/r \times C}$ and $\mathbf{W}_1 \in \mathbb{R}^{C \times C/r}$.

\textbf{Spatial Attention.} After channel refinement $\mathbf{F}' = \mathbf{M}_c(\mathbf{F}) \otimes \mathbf{F}$:
\begin{equation}
    \mathbf{M}_s(\mathbf{F}') = \sigma\!\Big(f^{7 \times 7}\!\big([\text{AvgPool}_c(\mathbf{F}') \,;\, \text{MaxPool}_c(\mathbf{F}')]\big)\Big)
\end{equation}
where $f^{7 \times 7}$ is a $7 \times 7$ convolution and $[\cdot\,;\,\cdot]$ denotes concatenation along the channel axis. The final output is:
\begin{equation}
    \mathbf{F}'' = \mathbf{M}_s(\mathbf{F}') \otimes \mathbf{F}'
\end{equation}

% ---- CBAM Diagram ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.4cm,
    box/.style={draw, rounded corners, minimum height=0.6cm, minimum width=3.0cm, align=center, font=\scriptsize},
    smallbox/.style={draw, rounded corners, minimum height=0.5cm, minimum width=1.3cm, align=center, font=\tiny},
    arrow/.style={-{Stealth[length=2mm]}, thick},
]
\node[box, fill=blue!10] (input) {Input Feature $\mathbf{F}$};
\node[smallbox, fill=yellow!15, below left=0.5cm and -0.2cm of input] (avgp1) {AvgPool};
\node[smallbox, fill=yellow!15, below right=0.5cm and -0.2cm of input] (maxp1) {MaxPool};
\node[box, fill=orange!12, below=1.2cm of input] (mlp) {Shared MLP ($C \to C/16 \to C$)};
\node[smallbox, fill=green!15, below=0.35cm of mlp] (sig1) {$\sigma$ (Sigmoid)};
\node[box, fill=red!8, below=0.35cm of sig1] (chanout) {$\mathbf{F}' = \mathbf{M}_c \otimes \mathbf{F}$};
\node[smallbox, fill=yellow!15, below left=0.5cm and -0.2cm of chanout] (avgp2) {AvgPool$_c$};
\node[smallbox, fill=yellow!15, below right=0.5cm and -0.2cm of chanout] (maxp2) {MaxPool$_c$};
\node[box, fill=orange!12, below=1.2cm of chanout] (conv) {$7 \times 7$ Conv};
\node[smallbox, fill=green!15, below=0.35cm of conv] (sig2) {$\sigma$ (Sigmoid)};
\node[box, fill=cyan!12, below=0.35cm of sig2] (output) {$\mathbf{F}'' = \mathbf{M}_s \otimes \mathbf{F}'$};

\draw[arrow] (input) -- (avgp1);
\draw[arrow] (input) -- (maxp1);
\draw[arrow] (avgp1) -- (mlp);
\draw[arrow] (maxp1) -- (mlp);
\draw[arrow] (mlp) -- (sig1);
\draw[arrow] (sig1) -- (chanout);
\draw[arrow] (chanout) -- (avgp2);
\draw[arrow] (chanout) -- (maxp2);
\draw[arrow] (avgp2) -- (conv);
\draw[arrow] (maxp2) -- (conv);
\draw[arrow] (conv) -- (sig2);
\draw[arrow] (sig2) -- (output);

\node[font=\tiny\bfseries, gray, left=0.1cm of mlp] {Channel Attn};
\node[font=\tiny\bfseries, gray, left=0.1cm of conv] {Spatial Attn};
\end{tikzpicture}
\caption{CBAM attention module. Channel attention learns \textit{which} features matter; spatial attention learns \textit{where} to focus. Total cost: 11,402 parameters across 5 decoder blocks.}
\label{fig:cbam}
\end{figure}

\subsection{Loss Functions}

\subsubsection{Binary Cross-Entropy (BCE)}

The pixel-level BCE loss measures the log-likelihood of each pixel prediction:
\begin{equation}
    \mathcal{L}_{\text{BCE}} = -\frac{1}{N}\sum_{i=1}^{N} \Big[ m_i \log(\hat{m}_i) + (1 - m_i) \log(1 - \hat{m}_i) \Big]
\end{equation}
where $m_i \in \{0,1\}$ is the ground truth label for pixel $i$, $\hat{m}_i \in (0,1)$ is the predicted probability, and $N = H \times W$ is the total number of pixels.

We use Soft BCE with label smoothing (SMP's \texttt{SoftBCEWithLogitsLoss}) which replaces hard targets with smoothed values $m_i' = m_i(1 - \alpha) + \alpha/2$ where $\alpha$ is the smoothing factor.

\subsubsection{Dice Loss}

The Dice loss~\cite{milletari2016vnet} directly optimizes the overlap between predicted and ground truth masks:
\begin{equation}
    \mathcal{L}_{\text{Dice}} = 1 - \frac{2\sum_{i=1}^{N} \hat{m}_i \cdot m_i + \epsilon}{\sum_{i=1}^{N} \hat{m}_i + \sum_{i=1}^{N} m_i + \epsilon}
\end{equation}

This is particularly important for our task because the Dice coefficient naturally handles class imbalance~--- it measures the ratio of overlap to total area, so a small tampered region is weighted proportionally to its size rather than being overwhelmed by the much larger authentic region.

\subsubsection{Combined Loss (Baseline)}

The total training loss is:
\begin{equation}
    \mathcal{L}_{\text{total}} = \mathcal{L}_{\text{BCE}} + \mathcal{L}_{\text{Dice}}
    \label{eq:loss}
\end{equation}

\subsubsection{Focal Loss (Ablation Variant)}

The Focal Loss~\cite{lin2017focal}, tested in experiment P.9, down-weights easy examples:
\begin{equation}
    \mathcal{L}_{\text{Focal}} = -\frac{1}{N}\sum_{i=1}^{N} \alpha_t (1 - p_t)^\gamma \log(p_t)
\end{equation}
where $p_t = \hat{m}_i$ if $m_i = 1$ and $p_t = 1 - \hat{m}_i$ otherwise, $\alpha = 0.25$, and $\gamma = 2.0$. The Focal+Dice variant (P.9) proved essentially neutral (+0.03pp), confirming that the Dice component already handles class imbalance sufficiently.

\subsection{Evaluation Metrics}

All segmentation metrics are computed on \textbf{tampered images only} (from vR.P.3 onward), since authentic images with all-zero masks trivially score perfect F1/IoU.

\subsubsection{Pixel-Level Metrics}

\textbf{Pixel F1 Score} (primary metric):
\begin{equation}
    F_1 = \frac{2 \cdot \text{TP}}{2 \cdot \text{TP} + \text{FP} + \text{FN}}
\end{equation}
where TP, FP, FN are pixel counts after binarizing predictions at $\tau = 0.5$.

\textbf{Intersection over Union (IoU / Jaccard Index):}
\begin{equation}
    \text{IoU} = \frac{\text{TP}}{\text{TP} + \text{FP} + \text{FN}} = \frac{|\hat{M} \cap M|}{|\hat{M} \cup M|}
\end{equation}

\textbf{Pixel AUC:} Area under the ROC curve at the pixel level, computed over all threshold values. This is threshold-independent and measures the model's discrimination quality:
\begin{equation}
    \text{AUC} = \int_0^1 \text{TPR}(\text{FPR}^{-1}(t))\, dt
\end{equation}

\textbf{Pixel Precision and Recall:}
\begin{equation}
    P = \frac{\text{TP}}{\text{TP} + \text{FP}}, \quad R = \frac{\text{TP}}{\text{TP} + \text{FN}}
\end{equation}

\subsubsection{Image-Level Metrics}

Image-level classification is derived from the mask: if any pixel exceeds $\tau$, the image is classified as tampered. We report Image Accuracy, Image Macro F1, and Image ROC-AUC.

\subsection{Training Configuration}

\begin{table}[H]
\centering
\caption{Training Hyperparameters}
\label{tab:training_config}
\begin{tabular}{l l}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Optimizer & Adam~\cite{kingma2015adam} ($\beta_1\!=\!0.9$, $\beta_2\!=\!0.999$) \\
Weight decay & $10^{-5}$ \\
Learning rate & $10^{-3}$ (single LR) \\
Loss function & BCE + Dice~\cite{milletari2016vnet} \\
LR scheduler & ReduceLROnPlateau \\
\quad Factor / Patience & 0.5 / 3 epochs \\
Early stopping & Patience 7 on val loss \\
Batch size & 16 \\
Max epochs & 25 (50 for extended) \\
Mixed precision & AMP~\cite{micikevicius2018mixed} + TF32 \\
Hardware & Kaggle T4 / P100 GPUs \\
Random seed & 42 \\
Framework & PyTorch + SMP~\cite{yakubovskiy2019smp} \\
\bottomrule
\end{tabular}
\end{table}

The Adam optimizer~\cite{kingma2015adam} update rule for parameter $\theta$ at step $t$:
\begin{align}
    m_t &= \beta_1 m_{t-1} + (1 - \beta_1) g_t \\
    v_t &= \beta_2 v_{t-1} + (1 - \beta_2) g_t^2 \\
    \theta_t &= \theta_{t-1} - \frac{\eta}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t
\end{align}
where $g_t = \nabla_\theta \mathcal{L}$, $\hat{m}_t = m_t/(1-\beta_1^t)$, $\hat{v}_t = v_t/(1-\beta_2^t)$.

\section{Research Journey: Four Experimental Lineages}

This project did not follow a straight line. Each failed approach taught lessons that directly shaped the next attempt. We document all four tracks because the failures were as informative as the successes.

\subsection{Track 1: Documentation-First (v0x) --- Failed}

The project began with an extensive literature review. A survey of 21 research papers was conducted, and documentation was written before any code: architecture specifications, dataset analyses, training plans~--- all generated with AI assistance (Gemini Deep Research, ChatGPT, Codex). The philosophy was ``get the docs right, then build notebooks from them.''

This approach failed when too many ideas from research papers were added simultaneously, leading to untested complexity. There was little to no executable code until late versions.

\textbf{Lesson:} Documentation without experimentation leads to untested assumptions. Ideas must be validated incrementally.

\subsection{Track 2: Kaggle Notebook Reproduction (vK) --- Failed}

A Kaggle notebook (\texttt{image-detection-with-mask}) was discovered that appeared to solve the assignment task: a dual-head UNet with classification and segmentation outputs, reporting $\sim$89.75\% accuracy and Dice $\sim$0.57 on CASIA.

An audit revealed two fatal problems:
\begin{enumerate}[nosep]
    \item \textbf{Data leakage:} The test CSV pointed to the validation set. Reported ``test accuracy'' was actually validation accuracy.
    \item \textbf{Segmentation failure:} On a dataset where $\sim$40\% of images are authentic (all-zero masks), a model predicting all-zero masks scores Dice $\sim$0.58. The segmentation head was not learning meaningful features.
\end{enumerate}

After fixing the leakage, training from scratch on $\sim$10K RGB images with 31.6M parameters proved insufficient. Best honest Tampered F1 in this track: 0.4101 (v6.5).

\textbf{Lesson:} Always audit for data leakage. Training from scratch on small datasets with large models is insufficient~--- the model memorizes low-level features instead of learning forensic patterns.

\subsection{Track 3: ETASR Classification (vR) --- Archived}

The project pivoted to reproducing a published research paper~\cite{nagm2024etasr}: ``Enhanced Image Tampering Detection using ELA and a CNN'' which reported 96.21\% accuracy using a custom 2-layer CNN on $128 \times 128$ ELA images.

Seven honest ablation experiments (vR.1.1 through vR.1.7) were conducted:

\begin{table}[H]
\centering
\caption{ETASR Classification Track Results}
\label{tab:etasr}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l l c c c l}
\toprule
\textbf{Ver.} & \textbf{Change} & \textbf{Acc} & \textbf{F1} & \textbf{Params} & \textbf{Verdict} \\
\midrule
vR.1.1 & Honest baseline & 88.38 & .8805 & 29.5M & Baseline \\
vR.1.2 & Data augmentation & 85.53 & .8505 & 29.5M & Rejected \\
vR.1.3 & Class weights & 89.17 & .8889 & 29.5M & Positive \\
vR.1.4 & BatchNorm & 88.75 & .8852 & 29.5M & Neutral \\
vR.1.5 & LR scheduler & 88.96 & .8873 & 29.5M & Neutral \\
\textbf{vR.1.6} & \textbf{Deeper CNN} & \textbf{90.23} & \textbf{.9004} & \textbf{13.8M} & \textbf{Best} \\
vR.1.7 & Global Avg Pool & 89.17 & .8901 & 64K & Neutral \\
\bottomrule
\end{tabular}
\end{table}

Key findings from this track:
\begin{itemize}[nosep]
    \item \textbf{Paper gap:} 5.98pp below the claimed 96.21\%, never closed (likely due to dataset filtering: the paper used JPEG-only images).
    \item \textbf{The Flatten-Dense bottleneck is toxic:} The Flatten layer connecting the final conv map to Dense(256) created a 29.5M-parameter bottleneck (99\%+ of total parameters). It caused data augmentation failure (vR.1.2) because Flatten memorizes exact pixel positions~--- augmented images activate completely different neurons.
    \item \textbf{Architecture $>$ training tricks:} The deeper CNN (vR.1.6, +1.85pp) outperformed class weights + BatchNorm + LR scheduler combined (+0.58pp).
    \item \textbf{Critical limitation:} Classification-only. No pixel-level localization capability.
\end{itemize}

The track was archived because the assignment requires localization.

\textbf{Lesson:} Verify the base architecture supports all required output tasks before investing in ablation studies.

\subsection{Track 4: Pretrained Ablation Study (vR.P) --- Final}

Armed with lessons from all prior failures, the final lineage used:
\begin{itemize}[nosep]
    \item \textbf{Pretrained encoder} (ResNet-34 on ImageNet) to avoid the small-dataset problem.
    \item \textbf{UNet segmentation} for pixel-level localization.
    \item \textbf{ELA preprocessing} (from ETASR track insight) as forensic input.
    \item \textbf{Single-variable ablation:} exactly one change per experiment.
    \item \textbf{Weights \& Biases}~\cite{wandb2020} tracking for all experiments.
    \item \textbf{Reproducibility protocol:} seed 42, identical hardware, two verification runs.
\end{itemize}

% ============================================================
% 6. COMPLETE ABLATION STUDY
% ============================================================
\section{Complete Ablation Study}

The ablation study is the core contribution of this work. Each experiment changed \textit{exactly one variable} from a known baseline, enabling clear cause-and-effect analysis.

\subsection{Full Results Table}

Table~\ref{tab:full_results} presents the complete results from all experiments in the vR.P track.

\begin{table*}[t]
\centering
\caption{Complete Experiment Results from the vR.P Ablation Study. Each row represents a single-variable change. Bold entries indicate the best value in each column. Verdicts: Strong+ = major improvement; Positive = meaningful improvement; Neutral = within noise; Negative = degradation.}
\label{tab:full_results}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l l c c c c c c c l}
\toprule
\textbf{Version} & \textbf{Modification} & \textbf{Px F1} & \textbf{IoU} & \textbf{Px AUC} & \textbf{Img Acc} & \textbf{Img F1} & \textbf{Img AUC} & \textbf{Params} & \textbf{Verdict} \\
\midrule
P.0 & Dataset baseline (no proper GT) & 0.3749 & 0.2307 & 0.8486 & 70.63\% & 0.6814 & 0.7860 & 3.17M & Baseline \\
P.1 & Proper GT masks & 0.4546 & 0.2942 & 0.8509 & 70.15\% & 0.6867 & 0.7785 & 3.17M & Baseline \\
P.1.5 & AMP + TF32 speed opts & 0.4227 & 0.2680 & 0.8560 & 71.05\% & 0.7016 & 0.7980 & 3.17M & Neutral \\
P.2 & Aggressive encoder unfreeze & 0.5117 & 0.3439 & 0.8688 & 69.04\% & 0.6673 & 0.7196 & 23.0M & Positive \\
\textbf{P.3} & \textbf{ELA input (replacing RGB)} & \textbf{0.6920} & 0.5291 & 0.9528 & 86.79\% & 0.8560 & 0.9502 & 3.17M & \textbf{Strong+} \\
P.4 & 4-channel RGB+ELA fusion & 0.7053 & 0.5447 & 0.9433 & 84.42\% & 0.8322 & 0.9229 & 3.17M & Neutral \\
P.5 & ResNet-50 encoder (RGB) & 0.5137 & 0.3456 & 0.8828 & 72.00\% & 0.7143 & 0.8126 & 9.0M & Positive \\
P.6 & EfficientNet-B0 encoder (RGB) & 0.5217 & 0.3529 & 0.8708 & 70.68\% & 0.6950 & 0.7801 & 2.24M & Positive \\
P.7 & Extended training (50 epochs) & 0.7154 & 0.5569 & 0.9504 & 87.37\% & 0.8637 & 0.9433 & 3.17M & Positive \\
P.8 & Progressive encoder unfreeze & 0.6985 & 0.5367 & 0.9541 & 87.59\% & 0.8650 & 0.9578 & 3.17M & Neutral \\
P.9 & Focal + Dice loss & 0.6923 & 0.5294 & 0.9323 & 87.16\% & 0.8606 & 0.9076 & 3.17M & Neutral \\
\textbf{P.10} & \textbf{CBAM attention in decoder} & 0.7277 & 0.5719 & 0.9573 & 87.32\% & 0.8615 & \textbf{0.9633} & 3.18M & \textbf{Positive} \\
P.12 & Data augmentation & 0.6968 & 0.5347 & 0.9502 & \textbf{88.48\%} & \textbf{0.8756} & 0.9427 & 3.17M & Neutral \\
P.14b & Test-time augmentation & 0.6388 & 0.4693 & \textbf{0.9618} & 87.43\% & 0.8619 & 0.9610 & 3.17M & Negative \\
\textbf{P.15} & \textbf{Multi-Q ELA (gray, Q=75/85/95)} & 0.7329 & 0.5785 & 0.9608 & 87.53\% & 0.8660 & 0.9423 & 3.17M & \textbf{Positive} \\
P.16 & DCT spatial map (replacing ELA) & 0.3209 & 0.1911 & 0.7778 & 61.60\% & 0.5678 & 0.6204 & 3.17M & Negative \\
P.17 & ELA + DCT fusion (6-channel) & 0.7302 & 0.5751 & 0.9431 & 87.06\% & 0.8589 & 0.9462 & 3.17M & Positive \\
\textbf{P.19} & \textbf{Multi-Q RGB ELA (9-channel)} & \textbf{0.7965} & \textbf{0.6615} & 0.9665 & --- & --- & --- & 3.17M & \textbf{Best} \\
P.27 & JPEG compression augmentation & 0.7523 & 0.6029 & --- & --- & --- & --- & 3.17M & Positive \\
P.28 & Cosine annealing LR scheduler & 0.7601 & 0.6132 & --- & --- & --- & --- & 3.17M & Positive \\
P.30 & Multi-Q ELA + CBAM (25ep) & 0.7714 & 0.6214 & --- & --- & --- & --- & 3.18M & Positive \\
P.30.1 & Multi-Q ELA + CBAM (50ep) & 0.7762 & 0.6344 & --- & --- & --- & --- & 3.18M & 2nd Best \\
P.30.2 & Multi-Q ELA + CBAM + unfreeze & 0.7721 & 0.6260 & --- & --- & --- & --- & 3.18M & Positive \\
P.30.3 & Multi-Q ELA + CBAM + focal & 0.7745 & --- & --- & --- & --- & --- & 3.18M & Positive \\
P.30.4 & Multi-Q ELA + CBAM + aug & 0.7745 & --- & --- & --- & --- & --- & 3.18M & Positive \\
\bottomrule
\end{tabular}
\end{table*}

\FloatBarrier

\subsection{Per-Metric Champions}

No single experiment dominates all metrics~--- a consistent finding across the study.

\begin{table}[H]
\centering
\caption{Per-Metric Champions}
\label{tab:champions}
\begin{tabular}{l l c}
\toprule
\textbf{Metric} & \textbf{Best Version} & \textbf{Value} \\
\midrule
\textbf{Pixel F1} & P.19 (Multi-Q RGB ELA) & \textbf{0.7965} \\
\textbf{Pixel IoU} & P.19 (Multi-Q RGB ELA) & \textbf{0.6615} \\
Pixel AUC & P.14b (TTA) & 0.9618 \\
Pixel Precision & P.8 (Prog. unfreeze) & 0.8857 \\
Image Accuracy & P.12 (Augmentation) & 88.48\% \\
Image Macro F1 & P.12 (Augmentation) & 0.8756 \\
Image AUC & P.10 (CBAM) & 0.9633 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablation by Category}

\subsubsection{Category 1: Input Representation (Dominant Factor)}

This was the most impactful category across the entire study. Table~\ref{tab:input_ablation} shows the input representation experiments.

\begin{table}[H]
\centering
\caption{Input Representation Ablation}
\label{tab:input_ablation}
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l c c c}
\toprule
\textbf{Input Type} & \textbf{Channels} & \textbf{Px F1} & \textbf{$\Delta$ from P.1} \\
\midrule
RGB (P.1) & 3 & 0.4546 & --- \\
ELA gray $\to$ RGB (P.3) & 3 & 0.6920 & +0.2374 \\
RGB + ELA (P.4) & 4 & 0.7053 & +0.2507 \\
Multi-Q ELA gray (P.15) & 3 & 0.7329 & +0.2783 \\
ELA + DCT (P.17) & 6 & 0.7302 & +0.2756 \\
\textbf{Multi-Q RGB ELA (P.19)} & \textbf{9} & \textbf{0.7965} & \textbf{+0.3419} \\
DCT only (P.16) & 3 & 0.3209 & $-$0.1337 \\
\bottomrule
\end{tabular}
\end{table}

Key observations:
\begin{itemize}[nosep]
    \item \textbf{RGB $\to$ ELA (P.3):} +23.74pp~--- the single largest improvement. False positive rate dropped from 22.6\% to 2.7\%.
    \item \textbf{Grayscale $\to$ RGB ELA (P.15 vs P.19):} Keeping RGB color information in each ELA map was worth +6.36pp over grayscale. The color channels carry additional forensic information about which color channels were manipulated.
    \item \textbf{DCT-only (P.16):} Catastrophic $-$37.11pp failure. Block-level frequency coefficients without spatial ELA context are insufficient for the ResNet encoder, which expects spatially structured input.
\end{itemize}

\subsubsection{Category 2: Encoder Architecture}

\begin{table}[H]
\centering
\caption{Encoder Architecture Comparison (all on RGB input)}
\label{tab:encoder_ablation}
\begin{tabular}{l c c c}
\toprule
\textbf{Encoder} & \textbf{Train. Params} & \textbf{Px F1} & \textbf{$\Delta$ from P.1} \\
\midrule
ResNet-34 (P.1) & 3.17M & 0.4546 & --- \\
ResNet-50 (P.5) & 9.0M & 0.5137 & +0.0591 \\
EfficientNet-B0 (P.6) & 2.24M & 0.5217 & +0.0671 \\
ResNet-34 + ELA (P.3) & 3.17M & \textbf{0.6920} & \textbf{+0.2374} \\
\bottomrule
\end{tabular}
\end{table}

The hierarchy is stark: \textbf{ELA on ResNet-34} (3.17M params, F1=0.6920) vastly outperforms ResNet-50 (9M params, F1=0.5137) and EfficientNet-B0 (2.24M params, F1=0.5217) on RGB. Input representation matters far more than encoder capacity.

\subsubsection{Category 3: Training Strategies}

\begin{table}[H]
\centering
\caption{Training Strategy Ablation (all from ELA baseline P.3)}
\label{tab:training_ablation}
\begin{tabular}{l c c}
\toprule
\textbf{Strategy} & \textbf{Px F1} & \textbf{$\Delta$ from P.3} \\
\midrule
P.3 Baseline (25 epochs) & 0.6920 & --- \\
P.7 Extended (50 epochs) & 0.7154 & +0.0234 \\
P.8 Progressive unfreeze & 0.6985 & +0.0065 \\
P.12 Data augmentation & 0.6968 & +0.0048 \\
P.1.5 AMP/TF32 & 0.4227 & (neutral) \\
\bottomrule
\end{tabular}
\end{table}

\begin{itemize}[nosep]
    \item \textbf{Extended training (P.7):} P.3's best epoch was 25/25 (the max). With 50-epoch budget, the optimum was found at epoch 36, yielding +2.34pp.
    \item \textbf{Progressive unfreeze (P.8):} Best epoch was in the frozen stage (epoch 23), before any unfreezing. Confirms frozen+BN is the sweet spot.
    \item \textbf{Data augmentation (P.12):} Only +0.48pp pixel F1 but achieved the best image accuracy (88.48\%) and image macro F1 (0.8756). Augmentation helps classification robustness more than pixel-level precision.
\end{itemize}

\subsubsection{Category 4: Loss Function}

Focal+Dice (P.9) was essentially neutral (+0.03pp) versus BCE+Dice. Pixel AUC actually regressed by $-$2.05pp because Focal Loss concentrates predictions near 0/1 extremes ($\gamma=2$ suppresses the gradient for well-classified pixels), destroying probability calibration.

\subsubsection{Category 5: Evaluation Strategy}

Test-time augmentation (P.14b) averaged predictions across 4 views (original, H-flip, V-flip, both flips). This achieved the best Pixel AUC (0.9618) and lowest FP rate (1.2\%) but \textit{destroyed} Pixel F1 ($-$5.32pp). Averaging smooths sharp mask boundaries, hurting localization precision.

\subsection{Impact Hierarchy}

\begin{table}[H]
\centering
\caption{Complete Impact Hierarchy (Delta from ELA baseline P.3)}
\label{tab:impact}
\begin{tabular}{c l l c}
\toprule
\textbf{\#} & \textbf{Category} & \textbf{Change} & \textbf{$\Delta$ F1 (pp)} \\
\midrule
1 & Input & Multi-Q RGB ELA 9ch & +10.45 \\
2 & Input & Multi-Q ELA gray & +4.09 \\
3 & Input & ELA+DCT fusion & +3.82 \\
4 & Architecture & CBAM attention & +3.57 \\
5 & Training & Extended training & +2.34 \\
6 & Input & RGB+ELA fusion & +1.33 \\
7 & Training & Progressive unfreeze & +0.65 \\
8 & Training & Augmentation & +0.48 \\
9 & Loss & Focal+Dice & +0.03 \\
10 & Evaluation & TTA & $-$5.32 \\
11 & Input & DCT-only & $-$37.11 \\
\bottomrule
\end{tabular}
\end{table}

The hierarchy is clear:
\begin{equation}
\boxed{\text{Input} \gg \text{Attention} > \text{Training} > \text{Loss}}
\end{equation}

\subsection{Parameter Efficiency Analysis}

CBAM provides the highest improvement density of any modification:
\begin{equation}
    \text{Efficiency}_{\text{CBAM}} = \frac{3.57\,\text{pp}}{11{,}402\,\text{params}} = 0.313\,\text{pp per 1K params}
\end{equation}

By contrast:
\begin{equation}
    \text{Efficiency}_{\text{ResNet-50}} = \frac{5.91\,\text{pp}}{5{,}830{,}000\,\text{extra params}} = 0.001\,\text{pp per 1K params}
\end{equation}

CBAM is $\sim$300$\times$ more parameter-efficient than upgrading the encoder.

\subsection{Diminishing Returns}

After Pixel F1 $\sim$0.78, combining improvements shows diminishing returns:
\begin{itemize}[nosep]
    \item P.19 alone: F1 = 0.7965
    \item P.30.1 (P.19 + CBAM + 50ep): F1 = 0.7762 ($<$ P.19)
    \item P.30.4 (P.19 + CBAM + aug): F1 = 0.7745 ($<$ P.19)
\end{itemize}

This suggests the model is approaching the ceiling of what the dataset and architecture can support. The 9-channel RGB ELA representation may already capture most of the forensic signal available in CASIA v2.0.

% ============================================================
% 7. ERROR ANALYSIS
% ============================================================
\section{Error Analysis}

\subsection{False Positive and False Negative Rates}

\begin{table}[H]
\centering
\caption{Image-Level Error Rates Across Configurations}
\label{tab:error_rates}
\begin{tabular}{l c c c}
\toprule
\textbf{Config} & \textbf{FP Rate} & \textbf{FN Rate} & \textbf{$\Delta$FP} \\
\midrule
RGB baseline (P.1) & 22.6\% & 40.4\% & --- \\
ELA baseline (P.3) & 2.7\% & 28.6\% & $-$19.9pp \\
CBAM (P.10) & 2.0\% & 28.3\% & $-$0.7pp \\
Augmentation (P.12) & 2.6\% & 24.6\% & $-$0.1pp \\
TTA (P.14b) & 1.2\% & 29.3\% & $-$1.5pp \\
\bottomrule
\end{tabular}
\end{table}

ELA reduces the false positive rate by 19.9 percentage points (22.6\% $\to$ 2.7\%). CBAM further drops it to 2.0\%. The remaining errors are primarily false negatives.

\subsection{Failure Modes}

The model struggles with three specific failure modes:
\begin{enumerate}[nosep]
    \item \textbf{Subtle copy-move forgeries:} When source and target textures are very similar, the ELA residuals are nearly identical, providing no forensic signal.
    \item \textbf{Very small tampered regions} ($<$2\% of image area): The 384$\times$384 resolution and the encoder's downsampling lose fine boundary details for tiny manipulations.
    \item \textbf{Multi-generation JPEG:} Images re-saved multiple times have uniform compression artifacts everywhere, destroying the differential ELA signal between authentic and tampered regions.
\end{enumerate}

\subsection{Precision-Recall Trade-off}

The best model (P.15 detailed metrics available) achieves:

\begin{table}[H]
\centering
\caption{Complete Metrics for Grayscale-ELA Model (P.15)}
\label{tab:detailed_metrics}
\begin{tabular}{l c l}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Interpretation} \\
\midrule
Pixel F1 & 0.7329 & Primary metric \\
Pixel IoU & 0.5785 & Stricter overlap \\
Pixel AUC & 0.9608 & Discrimination quality \\
Pixel Precision & 0.8409 & Few false alarms \\
Pixel Recall & 0.6496 & Misses some regions \\
Image Accuracy & 87.53\% & Detection accuracy \\
Image Macro F1 & 0.8660 & Balanced class perf \\
Image ROC-AUC & 0.9423 & Ranking quality \\
\bottomrule
\end{tabular}
\end{table}

The model is conservative: high precision (0.8409) means it rarely flags authentic regions as tampered, but moderate recall (0.6496) means it misses about 35\% of tampered pixels. This trade-off is acceptable for forensic applications where false accusations are costly.

% ============================================================
% EXPERIMENT LINEAGE DIAGRAM
% ============================================================
\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    node distance=0.55cm and 1.0cm,
    basenode/.style={draw, rounded corners=3pt, minimum height=0.7cm, minimum width=2.5cm, align=center, font=\scriptsize, thick},
    bestnode/.style={basenode, fill=green!25, line width=1.5pt},
    posnode/.style={basenode, fill=blue!10},
    negnode/.style={basenode, fill=red!15},
    neutnode/.style={basenode, fill=gray!10},
    mainflow/.style={-{Stealth[length=2.5mm]}, thick, black},
    branchflow/.style={-{Stealth[length=2mm]}, thick, black!60},
    delta/.style={font=\tiny, fill=white, inner sep=1pt, text=black!70},
]
% Main trunk
\node[neutnode] (p0) {P.0 RGB Baseline\\\textbf{F1 = 0.3749}};
\node[neutnode, below=0.5cm of p0] (p1) {P.1 Dataset Fix\\\textbf{F1 = 0.4546}};
\node[posnode, below=0.5cm of p1] (p3) {P.3 ELA Input\\\textbf{F1 = 0.6920}};
\node[posnode, below=0.5cm of p3] (p15) {P.15 Multi-Q ELA (gray)\\\textbf{F1 = 0.7329}};
\node[bestnode, below=0.5cm of p15] (p19) {P.19 Multi-Q RGB ELA 9ch\\\textbf{F1 = 0.7965}};
% Deltas on main trunk
\draw[mainflow] (p0) -- node[delta, right=2pt] {+0.0797} (p1);
\draw[mainflow] (p1) -- node[delta, right=2pt] {\textcolor{green!50!black}{+0.2374}} (p3);
\draw[mainflow] (p3) -- node[delta, right=2pt] {+0.0409} (p15);
\draw[mainflow] (p15) -- node[delta, right=2pt] {+0.0636} (p19);
% Branches from P.3
\node[posnode, right=1.8cm of p3] (p10) {P.10 CBAM Attention\\\textbf{F1 = 0.7277}};
\node[posnode, above=0.5cm of p10] (p7) {P.7 Extended 50ep\\\textbf{F1 = 0.7154}};
\node[negnode, below=0.5cm of p10] (p16) {P.16 DCT Only\\\textbf{F1 = 0.3209}};
\draw[branchflow] (p3.east) -- node[delta, above=1pt] {+0.0234} (p7.west);
\draw[branchflow] (p3.east) -- node[delta, above=1pt] {+0.0357} (p10.west);
\draw[branchflow] (p3.east) -- node[delta, below=1pt] {\textcolor{red!70!black}{$-$0.3711}} (p16.west);
% Branches from P.19
\node[posnode, right=1.8cm of p19] (p30) {P.30 + CBAM\\\textbf{F1 = 0.7714}};
\node[posnode, right=1.0cm of p30] (p301) {P.30.1 + 50ep\\\textbf{F1 = 0.7762}};
\draw[branchflow] (p19.east) -- (p30.west);
\draw[branchflow] (p30.east) -- (p301.west);
% Labels
\node[font=\scriptsize\bfseries, text=green!50!black, left=0.15cm of p19] {BEST};
% Legend
\node[anchor=north west, font=\scriptsize\bfseries] at ([xshift=4.5cm, yshift=0.2cm]p0.north east) (lt) {Legend:};
\node[bestnode, minimum width=1.1cm, minimum height=0.3cm, right=0.15cm of lt, font=\tiny] {Best};
\node[posnode, minimum width=1.1cm, minimum height=0.3cm, right=2.3cm of lt, font=\tiny] {Positive};
\node[negnode, minimum width=1.1cm, minimum height=0.3cm, right=4.3cm of lt, font=\tiny] {Negative};
\end{tikzpicture}
\caption{Experiment lineage diagram for the vR.P research track, showing evolution from RGB baseline (P.0) to best model (P.19). Delta annotations on arrows show Pixel F1 change. The largest single improvement (+0.2374) was switching from RGB to ELA.}
\label{fig:lineage}
\end{figure*}

\FloatBarrier


\section{Reproducibility and Experiment Tracking}

Reproducibility was a core design principle throughout this project.

\subsection{Fixed Experimental Constants}

To ensure valid single-variable comparisons, every experiment in the vR.P track shares:

\begin{table}[H]
\centering
\caption{Fixed Constants Across All vR.P Experiments}
\label{tab:constants}
\begin{tabular}{l l}
\toprule
\textbf{Constant} & \textbf{Value} \\
\midrule
Dataset & CASIA v2.0 (12,614 images) \\
Data split & 70/15/15 stratified \\
Random seed & 42 \\
Input resolution & 384$\times$384 \\
Base decoder & UNet (256/128/64/32/16) \\
Batch size & 16 \\
Optimizer & Adam ($\eta=10^{-3}$, wd=$10^{-5}$) \\
Framework & PyTorch 2.x + SMP \\
Hardware & Kaggle T4/P100 GPUs \\
Metric computation & Tampered images only \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Weights \& Biases Integration}

All experiments are logged to a centralized W\&B project~\cite{wandb2020} (\texttt{Tampered Image Detection \& Localization}), tracking:
\begin{itemize}[nosep]
    \item \textbf{Per-epoch:} \texttt{train\_loss}, \texttt{val\_loss}, \texttt{val\_pixel\_f1}, \texttt{val\_pixel\_iou}, \texttt{learning\_rate}
    \item \textbf{Final test:} \texttt{pixel\_f1}, \texttt{pixel\_iou}, \texttt{pixel\_auc}, \texttt{pixel\_precision}, \texttt{pixel\_recall}, \texttt{image\_accuracy}, \texttt{image\_macro\_f1}, \texttt{image\_roc\_auc}
    \item \textbf{Media:} \texttt{prediction\_examples} (W\&B Image panels)
\end{itemize}

The full interactive W\&B report is available at: \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}

\subsection{Infrastructure}

\begin{itemize}[nosep]
    \item \textbf{5 parallel Kaggle runner accounts} executing experiments via centralized configuration.
    \item \textbf{Papermill} for programmatic notebook execution from YAML experiment queues.
    \item \textbf{Centralized leaderboard notebook} that queries W\&B API and generates comparison tables, bar charts, and ablation progression plots automatically.
    \item \textbf{Google Colab T4 Optimised} notebooks for extended runs requiring more compute.
\end{itemize}

\subsection{Reproducibility Verification}

Two experiments were independently repeated to verify reproducibility:

\begin{table}[H]
\centering
\caption{Reproducibility Verification Results}
\label{tab:repro}
\begin{tabular}{l c c c}
\toprule
\textbf{Experiment} & \textbf{Run 1 F1} & \textbf{Run 2 F1} & \textbf{Match} \\
\midrule
vR.P.3 (ELA baseline) & 0.6920 & 0.6920 & Identical \\
vR.P.10 (CBAM) & 0.7277 & 0.7277 & Identical \\
\bottomrule
\end{tabular}
\end{table}

Both runs produced identical final metrics, confirming full reproducibility with seed 42 on Kaggle T4 GPUs.

% ============================================================
% 9. KEY LEARNINGS
% ============================================================
\section{Key Learnings}

The following lessons emerged from the full experimental journey across four research tracks and 60+ experiments:

\begin{enumerate}
    \item \textbf{Data leakage destroys credibility.} The Kaggle baseline (Track 2) appeared to work well ($\sim$89.75\% accuracy) but had inflated results due to test data leaking into training. Always audit data pipelines before trusting metrics.

    \item \textbf{Pretrained models are essential for small datasets.} Training from scratch on $\sim$10K images (Track 2) learns only low-level features and achieves Tampered F1 = 0.4101. A pretrained ResNet-34 on the same data achieves F1 = 0.4546, and with ELA reaches 0.6920.

    \item \textbf{Input representation matters more than everything else.} The ELA improvement (+23.74pp) exceeds all other improvements combined. Multi-quality RGB ELA adds another +10.45pp. No architectural change comes close.

    \item \textbf{Change one thing at a time.} Early attempts (Track 1) added multiple improvements simultaneously, leading to confusion and failure. Single-variable ablation is the only way to establish clear cause-and-effect relationships.

    \item \textbf{Negative results are valuable.} Knowing that DCT alone fails ($-$37.11pp), that TTA hurts localization ($-$5.32pp), and that Focal Loss is neutral (+0.03pp) prevents wasted effort in future work.

    \item \textbf{Verify the base approach supports all required tasks.} Track 3 (ETASR) achieved 90.23\% classification accuracy but was ultimately abandoned because it could not produce localization masks. This wasted significant effort.

    \item \textbf{Track experiments formally.} W\&B~\cite{wandb2020} made it possible to compare 60+ experiments systematically with reproducible metrics, automatic leaderboards, and visual comparisons. Without formal tracking, the project would have been chaotic.

    \item \textbf{The Flatten-Dense bottleneck is toxic.} In Track 3, the Flatten layer before Dense(256) created a 29.5M-parameter bottleneck that caused augmentation failure and training instability. Modern architectures avoid Flatten entirely using Global Average Pooling.

    \item \textbf{Freezing the encoder is the right default for small datasets.} Aggressive unfreezing (P.2, 23M trainable) overfit. Progressive unfreezing (P.8) confirmed that the best epoch was always in the frozen phase. Let the pretrained features do their job; only adapt BatchNorm.

    \item \textbf{Color information matters in forensic ELA.} Multi-Q RGB ELA (9ch, P.19, F1=0.7965) substantially outperformed Multi-Q grayscale ELA (3ch, P.15, F1=0.7329). The color channels encode which RGB channels were manipulated during forgery.
\end{enumerate}

% ============================================================
% 10. LIMITATIONS
% ============================================================
\section{Limitations}

\begin{enumerate}
    \item \textbf{Single dataset:} All results are on CASIA v2.0 only. Cross-dataset generalization to Columbia~\cite{hsu2006columbia}, Coverage~\cite{wen2016coverage}, CoMoFoD~\cite{tralic2013comofod}, and NIST~\cite{nist2019nimble} benchmarks is untested. Performance on out-of-distribution forgery types is unknown.

    \item \textbf{Fixed resolution:} 384$\times$384 input loses fine details, especially for small tampered regions. Higher resolution (512$\times$512 or 640$\times$640) may improve boundary precision but requires more GPU memory and longer training.

    \item \textbf{Binary segmentation only:} The model treats all tampering as one class. It does not distinguish between splicing and copy-move forgeries. Multi-class segmentation could provide more useful forensic information.

    \item \textbf{No robustness testing:} Behavior against post-processing attacks (JPEG recompression, Gaussian blur, noise addition, resizing) is not evaluated. Real-world tampered images often undergo multiple processing steps after manipulation.

    \item \textbf{No SRM noise features:} Spatial Rich Model~\cite{fridrich2012srm} noise extraction~--- a key technique in state-of-the-art systems like EMT-Net~\cite{emtnet2023} (AUC=0.987)~--- is not yet implemented.

    \item \textbf{ELA quality sensitivity:} The quality levels Q=75/85/95 were chosen by convention. A systematic quality sweep across all combinations was not performed.

    \item \textbf{Frozen encoder limitation:} Full fine-tuning with proper differential learning rates (lower LR for encoder, higher for decoder) and warmup scheduling is untested. This could potentially unlock better performance.

    \item \textbf{Diminishing returns ceiling:} After F1 $\sim$0.80, combining improvements shows diminishing or negative returns, suggesting the current architecture is approaching its ceiling on CASIA v2.0.
\end{enumerate}

% ============================================================
% 11. FUTURE WORK
% ============================================================
\section{Future Work}

\subsection{Immediate (High Confidence)}

\begin{itemize}[nosep]
    \item \textbf{Combine Multi-Q RGB ELA + CBAM with extended training:} The two best individual improvements have not been fully optimized together. The vR.P.30 series began this exploration, achieving F1=0.7762 vs.\ P.19's 0.7965, suggesting the combination needs different hyperparameters.
    \item \textbf{Differential learning rates:} Use lower LR ($10^{-5}$) for encoder BN layers and higher LR ($10^{-3}$) for decoder, with warmup scheduling.
    \item \textbf{Higher resolution input:} Test at 512$\times$512 and 640$\times$640 to improve boundary precision for small tampered regions.
\end{itemize}

\subsection{Medium-Term}

\begin{itemize}[nosep]
    \item \textbf{SRM noise filters:} Spatial Rich Model high-pass filters extract manipulation traces in the noise domain. Used by state-of-the-art systems.
    \item \textbf{YCbCr chrominance analysis:} Cb/Cr channels encode tamper traces better than luminance alone (reported 96.52\% on CASIA v2.0 with chrominance-only features).
    \item \textbf{JPEG quality augmentation:} Random re-compression during training to improve robustness against post-processing.
    \item \textbf{Cross-dataset evaluation:} Columbia~\cite{hsu2006columbia}, Coverage~\cite{wen2016coverage}, CoMoFoD~\cite{tralic2013comofod}, NIST~\cite{nist2019nimble} benchmarks to verify generalization.
\end{itemize}

\subsection{Long-Term}

\begin{itemize}[nosep]
    \item \textbf{Transformer-based encoders:} Swin Transformer or ConvNeXt may capture long-range manipulation patterns better than ResNet's local receptive fields.
    \item \textbf{Ensemble methods:} Combine predictions from Multi-Q ELA, CBAM, and augmentation variants using learned weights.
    \item \textbf{Multi-class segmentation:} Distinguish splicing from copy-move at the pixel level.
    \item \textbf{Robustness evaluation:} Systematic testing against JPEG recompression, blur, noise, and social media compression pipelines.
\end{itemize}

% ============================================================
% 12. CONCLUSION
% ============================================================
\section{Conclusion}

This report presented a systematic ablation study for image tamper detection and localization on the CASIA v2.0 dataset. Through 60+ controlled experiments across four research tracks, we demonstrated the following:

\begin{enumerate}
    \item \textbf{Input representation is the dominant factor.} Switching from raw RGB to Multi-Quality RGB ELA (9-channel) produced a +34.19 percentage point improvement in Pixel F1~--- far exceeding any architectural or training change.

    \item \textbf{The best model achieves Pixel F1 = 0.7965, IoU = 0.6615, and Pixel AUC = 0.9665}, using a UNet with a frozen pretrained ResNet-34 encoder and 9-channel multi-quality ELA input.

    \item \textbf{CBAM attention is the most parameter-efficient improvement}, adding +3.57pp Pixel F1 for only 11,402 parameters (0.36\% of trainable).

    \item \textbf{Single-variable ablation discipline} transforms ad-hoc experimentation into systematic scientific investigation, enabling clear cause-and-effect understanding.

    \item \textbf{Failures are as informative as successes.} Each failed research track (documentation-first, Kaggle reproduction, classification-only) contributed lessons that directly improved the final approach.
\end{enumerate}

The impact hierarchy established by this study:
\begin{equation}
\boxed{\text{Input representation} \gg \text{Attention} > \text{Training} > \text{Loss}}
\end{equation}
provides a clear roadmap for future work: invest in better input representations before trying more complex architectures.

All code, notebooks, and experiment logs are publicly available:
\begin{itemize}[nosep]
    \item GitHub: \url{https://github.com/The-Harsh-Vardhan/TIDAL-Tampered_Image_Detection_And_Localization}
    \item W\&B Report: \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}
\end{itemize}

% ============================================================
% REFERENCES
% ============================================================
\begin{thebibliography}{99}

\bibitem{dong2013casia}
J.~Dong, W.~Wang, and T.~Tan, ``CASIA Image Tampering Detection Evaluation Database,'' in \textit{Proc. IEEE China Summit and Int. Conf. on Signal and Information Processing}, pp.~422--426, 2013.

\bibitem{krawetz2007picture}
N.~Krawetz, ``A Picture's Worth\ldots Digital Image Analysis and Forensics,'' \textit{Hacker Factor Solutions}, 2007.

\bibitem{nagm2024etasr}
A.~Nagm \textit{et~al.}, ``Enhanced Image Tampering Detection using ELA and a CNN,'' \textit{Engineering, Technology \& Applied Science Research (ETASR)}, vol.~14, no.~1, pp.~12757--12762, 2024.

\bibitem{ronneberger2015unet}
O.~Ronneberger, P.~Fischer, and T.~Brox, ``U-Net: Convolutional Networks for Biomedical Image Segmentation,'' in \textit{Proc. MICCAI}, pp.~234--241, 2015.

\bibitem{he2016resnet}
K.~He, X.~Zhang, S.~Ren, and J.~Sun, ``Deep Residual Learning for Image Recognition,'' in \textit{Proc. IEEE CVPR}, pp.~770--778, 2016.

\bibitem{woo2018cbam}
S.~Woo, J.~Park, J.-Y.~Lee, and I.~S.~Kweon, ``CBAM: Convolutional Block Attention Module,'' in \textit{Proc. ECCV}, pp.~3--19, 2018.

\bibitem{yakubovskiy2019smp}
P.~Yakubovskiy, ``Segmentation Models PyTorch,'' GitHub, 2019. [Online]. Available: \url{https://github.com/qubvel/segmentation_models.pytorch}

\bibitem{bianchi2012jpeg}
T.~Bianchi and A.~Piva, ``Image Forgery Localization via Block-Grained Analysis of JPEG Artifacts,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~7, no.~3, pp.~1003--1017, 2012.

\bibitem{fridrich2012srm}
J.~Fridrich and J.~Kodovsky, ``Rich Models for Steganalysis of Digital Images,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~7, no.~3, pp.~868--882, 2012.

\bibitem{emtnet2023}
J.~Liu \textit{et~al.}, ``EMT-Net: Efficient Multi-Task Network for Image Manipulation Detection and Localization,'' in \textit{Proc. IEEE Int. Conf. on Multimedia and Expo (ICME)}, 2023.

\bibitem{deng2009imagenet}
J.~Deng, W.~Dong, R.~Socher, L.-J.~Li, K.~Li, and L.~Fei-Fei, ``ImageNet: A Large-Scale Hierarchical Image Database,'' in \textit{Proc. IEEE CVPR}, pp.~248--255, 2009.

\bibitem{milletari2016vnet}
F.~Milletari, N.~Navab, and S.-A.~Ahmadi, ``V-Net: Fully Convolutional Neural Networks for Volumetric Medical Image Segmentation,'' in \textit{Proc. Int. Conf. on 3D Vision (3DV)}, pp.~565--571, 2016.

\bibitem{lin2017focal}
T.-Y.~Lin, P.~Goyal, R.~Girshick, K.~He, and P.~Doll\'{a}r, ``Focal Loss for Dense Object Detection,'' in \textit{Proc. IEEE ICCV}, pp.~2980--2988, 2017.

\bibitem{kingma2015adam}
D.~P.~Kingma and J.~Ba, ``Adam: A Method for Stochastic Optimization,'' in \textit{Proc. ICLR}, 2015.

\bibitem{micikevicius2018mixed}
P.~Micikevicius \textit{et~al.}, ``Mixed Precision Training,'' in \textit{Proc. ICLR}, 2018.

\bibitem{zhou2018unetplusplus}
Z.~Zhou, M.~M.~Rahman~Siddiquee, N.~Tajbakhsh, and J.~Liang, ``UNet++: A Nested U-Net Architecture for Medical Image Segmentation,'' in \textit{Deep Learning in Medical Image Analysis}, pp.~3--11, 2018.

\bibitem{bayar2018constrained}
B.~Bayar and M.~C.~Stamm, ``Constrained Convolutional Neural Networks: A New Approach Towards General Purpose Image Manipulation Detection,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~13, no.~11, pp.~2691--2706, 2018.

\bibitem{wu2019mantranet}
Y.~Wu, W.~AbdAlmageed, and P.~Natarajan, ``ManTra-Net: Manipulation Tracing Network for Detection and Localization of Image Forgeries with Anomalous Features,'' in \textit{Proc. IEEE CVPR}, pp.~9543--9552, 2019.

\bibitem{wandb2020}
L.~Biewald, ``Experiment Tracking with Weights and Biases,'' 2020. [Online]. Available: \url{https://www.wandb.com}

\bibitem{hu2018senet}
J.~Hu, L.~Shen, and G.~Sun, ``Squeeze-and-Excitation Networks,'' in \textit{Proc. IEEE CVPR}, pp.~7132--7141, 2018.

\bibitem{hu2020span}
X.~Hu, Z.~Zhang, Z.~Jiang, S.~Chaudhuri, Z.~Yang, and R.~Nevatia, ``SPAN: Spatial Pyramid Attention Network for Image Manipulation Localization,'' in \textit{Proc. ECCV}, pp.~312--328, 2020.

\bibitem{wang2022objectformer}
J.~Wang, Z.~Wu, J.~Chen, X.~Han, A.~Shrivastava, S.~Lim, and Y.-G.~Jiang, ``ObjectFormer for Image Manipulation Detection and Localization,'' in \textit{Proc. IEEE CVPR}, pp.~2364--2373, 2022.

\bibitem{wu2022ifosn}
H.~Wu, J.~Zhou, J.~Tian, and J.~Liu, ``Robust Image Forgery Detection Against Transmission Over Online Social Networks,'' \textit{IEEE Trans. Information Forensics and Security}, vol.~17, pp.~443--456, 2022.

\bibitem{hsu2006columbia}
Y.-F.~Hsu and S.-F.~Chang, ``Detecting Image Splicing Using Geometry Invariants and Camera Characteristics Consistency,'' in \textit{Proc. IEEE Int. Conf. on Multimedia and Expo}, pp.~549--552, 2006.

\bibitem{wen2016coverage}
B.~Wen, Y.~Zhu, R.~Subramanian, T.-T.~Ng, X.~Shen, and S.~Winkler, ``COVERAGE~--- A Novel Database for Copy-Move Forgery Detection,'' in \textit{Proc. IEEE Int. Conf. on Image Processing (ICIP)}, pp.~161--165, 2016.

\bibitem{tralic2013comofod}
D.~Tralic, I.~Zupancic, S.~Grgic, and M.~Grgic, ``CoMoFoD~--- New Database for Copy-Move Forgery Detection,'' in \textit{Proc. ELMAR}, pp.~49--54, 2013.

\bibitem{nist2019nimble}
NIST, ``Media Forensics Challenge (MFC) and Nimble Challenge,'' National Institute of Standards and Technology, 2019.

\bibitem{tan2019efficientnet}
M.~Tan and Q.~V.~Le, ``EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks,'' in \textit{Proc. ICML}, pp.~6105--6114, 2019.

\bibitem{loshchilov2017cosine}
I.~Loshchilov and F.~Hutter, ``SGDR: Stochastic Gradient Descent with Warm Restarts,'' in \textit{Proc. ICLR}, 2017.

\bibitem{ioffe2015batchnorm}
S.~Ioffe and C.~Szegedy, ``Batch Normalization: Accelerating Deep Network Training by Reducing Internal Covariate Shift,'' in \textit{Proc. ICML}, pp.~448--456, 2015.

\end{thebibliography}

\end{document}
