\documentclass[11pt, twocolumn]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{float}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{placeins}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[all]{hypcap}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit, calc, decorations.pathreplacing, matrix}

\hypersetup{colorlinks=true, linkcolor=blue!70!black, citecolor=green!50!black, urlcolor=blue!60!black}

\title{\LARGE \textbf{Image Tampering Detection and Localization:\\A Systematic Ablation Study Using Deep Learning} \vspace{0.3cm} \\ \large Detailed Technical Report}

\author{
Harsh Vardhan \\
Big Vision Internship Assignment \\
\texttt{GitHub:} \url{https://github.com/The-Harsh-Vardhan/TIDAL-Tampered_Image_Detection_And_Localization} \\
\texttt{W\&B:} \url{https://api.wandb.ai/links/tampered-image-detection-and-localization/w5fy6lb0}
}

\date{}

\begin{document}
\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}
Digital images can be manipulated in ways invisible to the human eye. Detecting such tampering and pinpointing exactly which pixels were altered is a critical challenge in image forensics. This report presents a systematic ablation study for pixel-level image tamper detection and localization on the CASIA~v2.0 benchmark dataset (12,614 images: 7,491 authentic + 5,123 tampered). The project explored four experimental lineages over 60+ controlled experiments, evolving from documentation-first planning through Kaggle notebook reproduction, research paper baseline recreation, and finally a pretrained UNet segmentation pipeline.

The final system uses a UNet encoder-decoder with a frozen ResNet-34 encoder (ImageNet pretrained), multi-quality Error Level Analysis (ELA) preprocessing, and CBAM attention modules. Through strict single-variable ablation (one change per experiment), we established that \textbf{input representation dominates all other factors}~--- switching from RGB to multi-quality RGB ELA (9-channel) was worth +34.19 percentage points (pp) in Pixel~F1, more than all other improvements combined. The best model achieves \textbf{Pixel~F1 = 0.7965}, \textbf{IoU = 0.6615}, and \textbf{Pixel~AUC = 0.9665}. The impact hierarchy is clear: \textbf{Input representation $>$ Attention mechanisms $>$ Training configuration $>$ Loss function}.
\end{abstract}

\textbf{Keywords:} Image tampering detection, image forgery localization, error level analysis, UNet segmentation, ablation study, CASIA dataset, CBAM attention.

% ============================================================
% 1. INTRODUCTION
% ============================================================
\section{Introduction}

Digital images are ubiquitous in journalism, law enforcement, scientific research, and social media. However, modern image editing tools such as Adobe Photoshop, GIMP, and AI-based inpainting make it easy to manipulate photographs in ways that are nearly undetectable by the human eye. Common types of tampering include:

\begin{itemize}[nosep]
    \item \textbf{Splicing:} Copying a region from one image and pasting it into another.
    \item \textbf{Copy-move:} Duplicating a part of an image within the same image, often to hide or fabricate content.
    \item \textbf{Inpainting:} Using AI algorithms to fill in or replace regions of an image.
\end{itemize}

The task of \textit{image forgery detection} (classifying an image as authentic or tampered) is useful, but \textit{image forgery localization} (producing a pixel-level mask of tampered regions) is far more valuable for forensic analysis. This project tackles both tasks simultaneously: given an input image, the system produces a binary mask highlighting tampered pixels, from which image-level classification is derived.

\subsection{Problem Formulation}

Given an input image $\mathbf{X} \in \mathbb{R}^{H \times W \times 3}$, the goal is to learn a mapping function $f_\theta$ that produces a pixel-level probability map:
\begin{equation}
    \hat{\mathbf{M}} = f_\theta(\mathbf{X}) \in [0, 1]^{H \times W}
\end{equation}
where $\hat{\mathbf{M}}_{i,j}$ represents the probability that pixel $(i,j)$ has been tampered with. The ground truth mask $\mathbf{M} \in \{0, 1\}^{H \times W}$ is a binary map where 1 indicates tampered pixels and 0 indicates authentic pixels.

Image-level classification is derived from the segmentation output:
\begin{equation}
    \hat{y} = \begin{cases} 1 & \text{if } \max_{i,j} \hat{\mathbf{M}}_{i,j} > \tau \\ 0 & \text{otherwise} \end{cases}
\end{equation}
where $\tau = 0.5$ is the binarization threshold.

\subsection{Challenges}

Building an effective tamper localization system faces several fundamental difficulties:

\begin{enumerate}[nosep]
    \item \textbf{Extreme pixel-level class imbalance:} Tampered regions often occupy less than 2\% of all pixels in an image. A model that predicts all pixels as authentic achieves $>$98\% pixel accuracy while being completely useless.
    \item \textbf{Forgery sophistication:} Skilled forgers adjust lighting, color balance, and compression to make edits blend seamlessly with surrounding content.
    \item \textbf{JPEG compression artifacts:} Re-saving images as JPEG introduces compression artifacts that can either mask tampering traces or create false forensic signals.
    \item \textbf{Small dataset size:} Forensic datasets like CASIA v2.0 contain only $\sim$12K images~--- orders of magnitude smaller than ImageNet's 1.2M images~\cite{deng2009imagenet}, making training from scratch impractical.
\end{enumerate}

\subsection{Contributions}

The key contributions of this work are:
\begin{enumerate}[nosep]
    \item A complete pipeline from raw images to pixel-level tamper localization using multi-quality ELA preprocessing and pretrained UNet segmentation.
    \item A rigorous ablation study across 60+ experiments establishing that \textbf{input representation dominates all other factors}.
    \item Transparent documentation of four research tracks, including three failed approaches whose lessons directly shaped the final system.
    \item Full reproducibility via Weights~\&~Biases~\cite{wandb2020} experiment tracking, fixed random seeds, and open-source code.
\end{enumerate}

% ============================================================
% 2. RELATED WORK
% ============================================================
\section{Related Work}

\subsection{Traditional Forensic Methods}

Early forgery detection relied on hand-crafted features. Error Level Analysis (ELA)~\cite{krawetz2007picture} re-saves an image at a given JPEG quality and measures the pixel-wise difference from the original; tampered regions show inconsistent compression residuals. JPEG block artifact analysis~\cite{bianchi2012jpeg} exploits the 8$\times$8 block structure of JPEG compression to detect re-compressed regions. Noise-based methods analyze local noise variance~--- tampered regions composited from different sources often have mismatched noise patterns. Color Filter Array (CFA) analysis detects inconsistencies in the Bayer pattern interpolation. While interpretable, these methods struggle with sophisticated forgeries and produce high false-positive rates when used in isolation.

\subsection{CNN-Based Classification}

Deep learning transformed forgery detection by learning features directly from data. Bayar and Stamm~\cite{bayar2018constrained} proposed constrained convolutional layers that suppress image content and learn manipulation-specific features. Nagm et al.~\cite{nagm2024etasr} combined ELA preprocessing with a custom CNN for binary classification (tampered vs.\ authentic), reporting 96.21\% accuracy on CASIA v2.0 (JPEG-only subset). However, classification-only models cannot localize tampered regions~--- they answer ``is it tampered?'' but not ``where?''

\subsection{Segmentation-Based Localization}

To achieve pixel-level localization, researchers adopted encoder-decoder architectures. The UNet~\cite{ronneberger2015unet}, originally designed for biomedical image segmentation, proved effective for tamper localization through its skip connections that preserve fine spatial details. UNet++~\cite{zhou2018unetplusplus} improved boundary delineation through nested dense skip connections. Using pretrained encoders (e.g., ResNet~\cite{he2016resnet} trained on ImageNet~\cite{deng2009imagenet}) significantly improves performance on small forensic datasets by providing rich, generalizable feature representations.

ManTra-Net~\cite{wu2019mantranet} introduced a self-supervised anomaly detection approach for forgery localization that generalizes across manipulation types without per-dataset fine-tuning. SPAN~\cite{hu2020span} used spatial pyramid attention to capture multi-scale manipulation traces for more precise boundary localization.

\subsection{Attention Mechanisms}

Attention modules help models focus on the most relevant features. SE-Net~\cite{hu2018senet} introduced channel attention through squeeze-and-excitation blocks. CBAM~\cite{woo2018cbam} extended this with both channel-wise and spatial attention, teaching the model which feature maps and spatial locations are most important. In our experiments, CBAM provided one of the most parameter-efficient improvements (+3.57pp Pixel F1 for only 11K parameters).

\subsection{Multi-Stream and Frequency-Domain Methods}

State-of-the-art systems combine multiple input streams. EMT-Net~\cite{emtnet2023} uses Swin Transformers with RGB, noise maps, and frequency features, achieving AUC=0.987 on NIST. SRM (Spatial Rich Model)~\cite{fridrich2012srm} high-pass filters extract manipulation traces in the noise domain. ObjectFormer~\cite{wang2022objectformer} uses transformer-based architectures that learn object-level relationships between image patches to detect manipulated regions. IF-OSN~\cite{wu2022ifosn} combines frequency-domain analysis with online social network compression patterns for robust detection even after social media re-compression.

These multi-stream methods represent the current frontier but require significantly more computational resources and training data. Our work demonstrates that substantial performance can be achieved with simpler architectures when the input representation is well-designed.

% ============================================================
% 3. DATASET
% ============================================================
\section{Dataset}

\subsection{CASIA v2.0}

All experiments use the \textbf{CASIA v2.0} dataset~\cite{dong2013casia}, a widely-used benchmark for image forgery detection and localization.

\begin{table}[H]
\centering
\caption{CASIA v2.0 Dataset Summary}
\label{tab:dataset}
\begin{tabular}{l r}
\toprule
\textbf{Property} & \textbf{Value} \\
\midrule
Total images & 12,614 \\
Authentic images & 7,491 (59.4\%) \\
Tampered images & 5,123 (40.6\%) \\
\quad Copy-move forgeries & 3,295 (26.1\%) \\
\quad Splicing forgeries & 1,828 (14.5\%) \\
Ground truth & Binary masks \\
Image formats & JPEG, BMP, TIFF \\
Resolution range & Variable (resized to 384$\times$384) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Data Splitting}

The dataset is split using a two-step stratified process with seed 42:

\begin{enumerate}[nosep]
    \item 70/30 split of the full dataset (stratified by authentic/tampered label).
    \item 50/50 split of the 30\% holdout into validation and test sets.
\end{enumerate}

\begin{table}[H]
\centering
\caption{Data Split Configuration}
\label{tab:split}
\begin{tabular}{l r r l}
\toprule
\textbf{Split} & \textbf{Count} & \textbf{Ratio} & \textbf{Purpose} \\
\midrule
Training & 8,829 & 70\% & Model optimization \\
Validation & 1,892 & 15\% & Early stopping, LR scheduling \\
Test & 1,893 & 15\% & Final reported metrics \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Preprocessing}

\begin{itemize}[nosep]
    \item All images resized to $384 \times 384$ pixels using bilinear interpolation.
    \item Ground truth masks resized with nearest-neighbor interpolation, then binarized at threshold 0.5.
    \item Per-channel normalization statistics computed from 500 randomly sampled training images.
\end{itemize}

\subsection{Data Leakage Discovery}

An important finding during the project was that the Kaggle variant of the CASIA dataset mixed ground truth mask images into the input data folder. A CNN trivially distinguished mask images (simple binary patterns) from real photographs, producing 100\% ``test accuracy.'' This result was discarded, and the discovery motivated strict pipeline audits in all subsequent experiments.

% ============================================================
% 4. METHODOLOGY
% ============================================================
\section{Methodology}

The final system consists of three components: (1) multi-quality ELA preprocessing, (2) a UNet segmentation model with a pretrained frozen encoder and CBAM attention, and (3) a training pipeline with carefully chosen loss functions and hyperparameters.

% ---- Architecture Diagram (vertical, single-column) ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.6cm,
    box/.style={draw, rounded corners, minimum height=0.75cm, minimum width=5.6cm, align=center, font=\footnotesize},
    outbox/.style={draw, rounded corners, minimum height=0.7cm, minimum width=2.4cm, align=center, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
]
\node[box, fill=blue!10] (input) {Input Image (384$\times$384$\times$3 RGB)};
\node[box, fill=orange!15, below=of input] (ela) {Multi-Quality ELA Preprocessing\\Q=75, Q=85, Q=95 $\times$ RGB = 9 channels};
\node[box, fill=green!15, below=of ela] (encoder) {ResNet-34 Encoder\\(ImageNet Pretrained, Frozen + BN Unfrozen)};
\node[box, fill=purple!12, below=of encoder] (decoder) {UNet Decoder (5 blocks)\\+ CBAM Attention + Skip Connections};
\node[box, fill=yellow!12, below=of decoder] (sigmoid) {Sigmoid Output (384$\times$384$\times$1)};
\node[outbox, fill=cyan!15, below left=0.55cm and -0.3cm of sigmoid] (mask) {Segmentation\\Mask};
\node[outbox, fill=cyan!15, below right=0.55cm and -0.3cm of sigmoid] (label_out) {Image Label\\(Auth / Tampered)};
\draw[arrow] (input) -- (ela);
\draw[arrow] (ela) -- (encoder);
\draw[arrow] (encoder) -- (decoder);
\draw[arrow] (decoder) -- (sigmoid);
\draw[arrow] (sigmoid.south) -- ++(-1.3,-0.25) -- (mask.north);
\draw[arrow] (sigmoid.south) -- ++(1.3,-0.25) -- (label_out.north);
\draw[arrow, dashed, gray] (encoder.east) -- ++(0.3,0) |- node[pos=0.25, right, font=\scriptsize, gray, align=left] {skip\\connections} (decoder.east);
\node[font=\scriptsize, gray, right=0.05cm of label_out] {max pixel prob.};
\end{tikzpicture}
\caption{Overall architecture of the tamper detection and localization pipeline.}
\label{fig:architecture}
\end{figure}

\subsection{Error Level Analysis (ELA)}

ELA is a forensic preprocessing technique that reveals JPEG compression inconsistencies. The core idea is that authentic regions of an image compress uniformly, while tampered regions~--- composited from different sources or saved at different quality levels~--- show inconsistent compression residuals.

\subsubsection{Mathematical Formulation}

Given an input image $\mathbf{X} \in \mathbb{R}^{H \times W \times 3}$ and a JPEG quality level $Q \in \{1, \ldots, 100\}$, the ELA map is computed as:
\begin{equation}
    \text{ELA}_Q(\mathbf{X}) = \left| \mathbf{X} - \text{JPEG}_Q(\mathbf{X}) \right| \cdot s
    \label{eq:ela}
\end{equation}
where $\text{JPEG}_Q(\cdot)$ denotes JPEG re-compression at quality $Q$, and $s$ is a brightness scaling factor such that the maximum channel difference maps to 255:
\begin{equation}
    s = \frac{255}{\max_{i,j,c} \left| X_{i,j,c} - \text{JPEG}_Q(\mathbf{X})_{i,j,c} \right| + \epsilon}
\end{equation}
with $\epsilon = 10^{-8}$ for numerical stability.

\subsubsection{Multi-Quality RGB ELA (Best Variant)}

Instead of a single quality level, the best configuration uses three quality levels, each producing an RGB ELA map (3 channels). These are stacked to form a 9-channel input:
\begin{equation}
    \mathbf{X}_{\text{MQ-ELA}} = \left[ \text{ELA}_{75}(\mathbf{X}) \,\|\, \text{ELA}_{85}(\mathbf{X}) \,\|\, \text{ELA}_{95}(\mathbf{X}) \right] \in \mathbb{R}^{H \times W \times 9}
\end{equation}
where $\|$ denotes channel-wise concatenation.

Each quality level captures different forensic information:

\begin{table}[H]
\centering
\caption{Multi-Quality ELA Channel Statistics (from P.15 training set)}
\label{tab:ela_stats}
\begin{tabular}{l c c l}
\toprule
\textbf{Quality} & \textbf{Mean} & \textbf{Std} & \textbf{Sensitivity} \\
\midrule
Q=75 (aggressive) & 0.0684 & 0.0656 & Strong manipulations \\
Q=85 (balanced) & 0.0605 & 0.0604 & Medium edits \\
Q=95 (gentle) & 0.0402 & 0.0471 & Subtle changes \\
\bottomrule
\end{tabular}
\end{table}

The three channels carry meaningfully different statistics~--- lower $Q$ produces larger residuals, confirming they provide distinct, non-redundant forensic information.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.45cm,
    box/.style={draw, rounded corners, minimum height=0.65cm, minimum width=2.6cm, align=center, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
]
\node[box, fill=orange!10] (q75) {Re-save at Q=75};
\node[box, fill=orange!10, right=0.25cm of q75] (q85) {Re-save at Q=85};
\node[box, fill=orange!10, right=0.25cm of q85] (q95) {Re-save at Q=95};
\node[box, fill=blue!10, above=0.5cm of q85] (orig) {Original Image (RGB)};
\node[box, fill=yellow!15, below=0.5cm of q85] (diff) {$|\text{Original} - \text{Recompressed}|$};
\node[box, fill=green!15, below=0.4cm of diff] (stack) {Stack $\rightarrow$ 9-Channel Tensor};
\node[box, fill=purple!10, below=0.4cm of stack] (model) {Feed to UNet Model};
\draw[arrow] (orig) -- (q75);
\draw[arrow] (orig) -- (q85);
\draw[arrow] (orig) -- (q95);
\draw[arrow] (q75) -- (diff);
\draw[arrow] (q85) -- (diff);
\draw[arrow] (q95) -- (diff);
\draw[arrow] (diff) -- (stack);
\draw[arrow] (stack) -- (model);
\end{tikzpicture}
\caption{Multi-Quality RGB ELA preprocessing pipeline.}
\label{fig:ela_pipeline}
\end{figure}

\subsubsection{ELA Algorithm}

\begin{algorithm}[H]
\caption{Multi-Quality RGB ELA Computation}
\label{alg:ela}
\begin{algorithmic}[1]
\Require Input image $\mathbf{X}$, quality levels $\mathcal{Q} = \{75, 85, 95\}$
\Ensure 9-channel ELA tensor $\mathbf{X}_{\text{MQ-ELA}}$
\State $\text{channels} \gets [\,]$
\For{$Q \in \mathcal{Q}$}
    \State $\mathbf{X}_Q \gets \text{JPEG\_recompress}(\mathbf{X}, Q)$
    \State $\mathbf{D}_Q \gets |\mathbf{X} - \mathbf{X}_Q|$
    \State $s_Q \gets 255 \,/\, (\max(\mathbf{D}_Q) + \epsilon)$
    \State $\text{channels.append}(\mathbf{D}_Q \cdot s_Q)$ \Comment{3 RGB channels}
\EndFor
\State $\mathbf{X}_{\text{MQ-ELA}} \gets \text{concatenate}(\text{channels}, \text{axis}=\text{channel})$
\State \Return $\mathbf{X}_{\text{MQ-ELA}}$
\end{algorithmic}
\end{algorithm}

\subsection{Model Architecture}

\subsubsection{UNet with ResNet-34 Encoder}

The model uses the Segmentation Models PyTorch (SMP) library's~\cite{yakubovskiy2019smp} UNet implementation with a ResNet-34~\cite{he2016resnet} encoder.

The ResNet-34 encoder uses residual connections:
\begin{equation}
    \mathbf{h}_l = \sigma(\mathbf{W}_l \ast \mathbf{h}_{l-1} + \mathbf{h}_{l-1})
\end{equation}
where $\ast$ denotes convolution, $\sigma$ is the activation function (ReLU), and the identity shortcut $\mathbf{h}_{l-1}$ allows gradients to flow directly through the network.

\textbf{Freeze strategy:} All encoder convolutional weights are frozen. Only BatchNorm layers remain unfrozen (trainable scale $\gamma$ and shift $\beta$ parameters). This allows BN statistics to adapt from ImageNet's natural image distribution to the ELA forensic domain:
\begin{equation}
    \hat{x} = \gamma \cdot \frac{x - \mu_{\text{batch}}}{\sqrt{\sigma^2_{\text{batch}} + \epsilon}} + \beta
\end{equation}

The first convolutional layer is modified to accept 9 input channels (from 3) by repeating the pretrained weights 3 times.

\begin{table}[H]
\centering
\caption{UNet Decoder Configuration}
\label{tab:decoder}
\begin{tabular}{c c c c}
\toprule
\textbf{Block} & \textbf{Channels} & \textbf{Input Scale} & \textbf{Skip From} \\
\midrule
1 & 256 & 1/16 & Encoder stage 4 \\
2 & 128 & 1/8 & Encoder stage 3 \\
3 & 64 & 1/4 & Encoder stage 2 \\
4 & 32 & 1/2 & Encoder stage 1 \\
5 & 16 & 1/1 & Input \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Model Parameter Summary}
\label{tab:params}
\begin{tabular}{l r r}
\toprule
\textbf{Component} & \textbf{Parameters} & \textbf{Trainable} \\
\midrule
Encoder (ResNet-34) & 21.3M & BN only \\
Decoder (5 blocks) & $\sim$3.17M & Yes \\
CBAM modules & 11,402 & Yes \\
Segmentation head & $\sim$1K & Yes \\
\midrule
\textbf{Total} & \textbf{24.4M} & \textbf{3.17M (13\%)} \\
\bottomrule
\end{tabular}
\end{table}

This freeze strategy was validated empirically: P.2's aggressive unfreeze (23M trainable) overfit with a 1:2,615 data-to-parameter ratio. P.8's progressive unfreeze confirmed the best epoch occurred before any unfreezing stage.

\subsubsection{CBAM Attention Module}

CBAM~\cite{woo2018cbam} is injected into all 5 UNet decoder blocks. It applies two sequential attention operations:

\textbf{Channel Attention.} Given a feature map $\mathbf{F} \in \mathbb{R}^{C \times H \times W}$:
\begin{equation}
    \mathbf{M}_c(\mathbf{F}) = \sigma\!\Big(\text{MLP}\!\big(\text{AvgPool}(\mathbf{F})\big) + \text{MLP}\!\big(\text{MaxPool}(\mathbf{F})\big)\Big)
\end{equation}
where $\sigma$ is the sigmoid function and MLP is a shared two-layer network with bottleneck ratio $r=16$:
\begin{equation}
    \text{MLP}(\mathbf{z}) = \mathbf{W}_1 \cdot \text{ReLU}(\mathbf{W}_0 \cdot \mathbf{z})
\end{equation}
with $\mathbf{W}_0 \in \mathbb{R}^{C/r \times C}$ and $\mathbf{W}_1 \in \mathbb{R}^{C \times C/r}$.

\textbf{Spatial Attention.} After channel refinement $\mathbf{F}' = \mathbf{M}_c(\mathbf{F}) \otimes \mathbf{F}$:
\begin{equation}
    \mathbf{M}_s(\mathbf{F}') = \sigma\!\Big(f^{7 \times 7}\!\big([\text{AvgPool}_c(\mathbf{F}') \,;\, \text{MaxPool}_c(\mathbf{F}')]\big)\Big)
\end{equation}
where $f^{7 \times 7}$ is a $7 \times 7$ convolution and $[\cdot\,;\,\cdot]$ denotes concatenation along the channel axis. The final output is:
\begin{equation}
    \mathbf{F}'' = \mathbf{M}_s(\mathbf{F}') \otimes \mathbf{F}'
\end{equation}

% ---- CBAM Diagram ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.4cm,
    box/.style={draw, rounded corners, minimum height=0.6cm, minimum width=3.0cm, align=center, font=\scriptsize},
    smallbox/.style={draw, rounded corners, minimum height=0.5cm, minimum width=1.3cm, align=center, font=\tiny},
    arrow/.style={-{Stealth[length=2mm]}, thick},
]
\node[box, fill=blue!10] (input) {Input Feature $\mathbf{F}$};
\node[smallbox, fill=yellow!15, below left=0.5cm and -0.2cm of input] (avgp1) {AvgPool};
\node[smallbox, fill=yellow!15, below right=0.5cm and -0.2cm of input] (maxp1) {MaxPool};
\node[box, fill=orange!12, below=1.2cm of input] (mlp) {Shared MLP ($C \to C/16 \to C$)};
\node[smallbox, fill=green!15, below=0.35cm of mlp] (sig1) {$\sigma$ (Sigmoid)};
\node[box, fill=red!8, below=0.35cm of sig1] (chanout) {$\mathbf{F}' = \mathbf{M}_c \otimes \mathbf{F}$};
\node[smallbox, fill=yellow!15, below left=0.5cm and -0.2cm of chanout] (avgp2) {AvgPool$_c$};
\node[smallbox, fill=yellow!15, below right=0.5cm and -0.2cm of chanout] (maxp2) {MaxPool$_c$};
\node[box, fill=orange!12, below=1.2cm of chanout] (conv) {$7 \times 7$ Conv};
\node[smallbox, fill=green!15, below=0.35cm of conv] (sig2) {$\sigma$ (Sigmoid)};
\node[box, fill=cyan!12, below=0.35cm of sig2] (output) {$\mathbf{F}'' = \mathbf{M}_s \otimes \mathbf{F}'$};

\draw[arrow] (input) -- (avgp1);
\draw[arrow] (input) -- (maxp1);
\draw[arrow] (avgp1) -- (mlp);
\draw[arrow] (maxp1) -- (mlp);
\draw[arrow] (mlp) -- (sig1);
\draw[arrow] (sig1) -- (chanout);
\draw[arrow] (chanout) -- (avgp2);
\draw[arrow] (chanout) -- (maxp2);
\draw[arrow] (avgp2) -- (conv);
\draw[arrow] (maxp2) -- (conv);
\draw[arrow] (conv) -- (sig2);
\draw[arrow] (sig2) -- (output);

\node[font=\tiny\bfseries, gray, left=0.1cm of mlp] {Channel Attn};
\node[font=\tiny\bfseries, gray, left=0.1cm of conv] {Spatial Attn};
\end{tikzpicture}
\caption{CBAM attention module. Channel attention learns \textit{which} features matter; spatial attention learns \textit{where} to focus. Total cost: 11,402 parameters across 5 decoder blocks.}
\label{fig:cbam}
\end{figure}

\subsection{Loss Functions}

\subsubsection{Binary Cross-Entropy (BCE)}

The pixel-level BCE loss measures the log-likelihood of each pixel prediction:
\begin{equation}
    \mathcal{L}_{\text{BCE}} = -\frac{1}{N}\sum_{i=1}^{N} \Big[ m_i \log(\hat{m}_i) + (1 - m_i) \log(1 - \hat{m}_i) \Big]
\end{equation}
where $m_i \in \{0,1\}$ is the ground truth label for pixel $i$, $\hat{m}_i \in (0,1)$ is the predicted probability, and $N = H \times W$ is the total number of pixels.

We use Soft BCE with label smoothing (SMP's \texttt{SoftBCEWithLogitsLoss}) which replaces hard targets with smoothed values $m_i' = m_i(1 - \alpha) + \alpha/2$ where $\alpha$ is the smoothing factor.

\subsubsection{Dice Loss}

The Dice loss~\cite{milletari2016vnet} directly optimizes the overlap between predicted and ground truth masks:
\begin{equation}
    \mathcal{L}_{\text{Dice}} = 1 - \frac{2\sum_{i=1}^{N} \hat{m}_i \cdot m_i + \epsilon}{\sum_{i=1}^{N} \hat{m}_i + \sum_{i=1}^{N} m_i + \epsilon}
\end{equation}

This is particularly important for our task because the Dice coefficient naturally handles class imbalance~--- it measures the ratio of overlap to total area, so a small tampered region is weighted proportionally to its size rather than being overwhelmed by the much larger authentic region.

\subsubsection{Combined Loss (Baseline)}

The total training loss is:
\begin{equation}
    \mathcal{L}_{\text{total}} = \mathcal{L}_{\text{BCE}} + \mathcal{L}_{\text{Dice}}
    \label{eq:loss}
\end{equation}

\subsubsection{Focal Loss (Ablation Variant)}

The Focal Loss~\cite{lin2017focal}, tested in experiment P.9, down-weights easy examples:
\begin{equation}
    \mathcal{L}_{\text{Focal}} = -\frac{1}{N}\sum_{i=1}^{N} \alpha_t (1 - p_t)^\gamma \log(p_t)
\end{equation}
where $p_t = \hat{m}_i$ if $m_i = 1$ and $p_t = 1 - \hat{m}_i$ otherwise, $\alpha = 0.25$, and $\gamma = 2.0$. The Focal+Dice variant (P.9) proved essentially neutral (+0.03pp), confirming that the Dice component already handles class imbalance sufficiently.

\subsection{Evaluation Metrics}

All segmentation metrics are computed on \textbf{tampered images only} (from vR.P.3 onward), since authentic images with all-zero masks trivially score perfect F1/IoU.

\subsubsection{Pixel-Level Metrics}

\textbf{Pixel F1 Score} (primary metric):
\begin{equation}
    F_1 = \frac{2 \cdot \text{TP}}{2 \cdot \text{TP} + \text{FP} + \text{FN}}
\end{equation}
where TP, FP, FN are pixel counts after binarizing predictions at $\tau = 0.5$.

\textbf{Intersection over Union (IoU / Jaccard Index):}
\begin{equation}
    \text{IoU} = \frac{\text{TP}}{\text{TP} + \text{FP} + \text{FN}} = \frac{|\hat{M} \cap M|}{|\hat{M} \cup M|}
\end{equation}

\textbf{Pixel AUC:} Area under the ROC curve at the pixel level, computed over all threshold values. This is threshold-independent and measures the model's discrimination quality:
\begin{equation}
    \text{AUC} = \int_0^1 \text{TPR}(\text{FPR}^{-1}(t))\, dt
\end{equation}

\textbf{Pixel Precision and Recall:}
\begin{equation}
    P = \frac{\text{TP}}{\text{TP} + \text{FP}}, \quad R = \frac{\text{TP}}{\text{TP} + \text{FN}}
\end{equation}

\subsubsection{Image-Level Metrics}

Image-level classification is derived from the mask: if any pixel exceeds $\tau$, the image is classified as tampered. We report Image Accuracy, Image Macro F1, and Image ROC-AUC.

\subsection{Training Configuration}

\begin{table}[H]
\centering
\caption{Training Hyperparameters}
\label{tab:training_config}
\begin{tabular}{l l}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Optimizer & Adam~\cite{kingma2015adam} ($\beta_1\!=\!0.9$, $\beta_2\!=\!0.999$) \\
Weight decay & $10^{-5}$ \\
Learning rate & $10^{-3}$ (single LR) \\
Loss function & BCE + Dice~\cite{milletari2016vnet} \\
LR scheduler & ReduceLROnPlateau \\
\quad Factor / Patience & 0.5 / 3 epochs \\
Early stopping & Patience 7 on val loss \\
Batch size & 16 \\
Max epochs & 25 (50 for extended) \\
Mixed precision & AMP~\cite{micikevicius2018mixed} + TF32 \\
Hardware & Kaggle T4 / P100 GPUs \\
Random seed & 42 \\
Framework & PyTorch + SMP~\cite{yakubovskiy2019smp} \\
\bottomrule
\end{tabular}
\end{table}

The Adam optimizer~\cite{kingma2015adam} update rule for parameter $\theta$ at step $t$:
\begin{align}
    m_t &= \beta_1 m_{t-1} + (1 - \beta_1) g_t \\
    v_t &= \beta_2 v_{t-1} + (1 - \beta_2) g_t^2 \\
    \theta_t &= \theta_{t-1} - \frac{\eta}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t
\end{align}
where $g_t = \nabla_\theta \mathcal{L}$, $\hat{m}_t = m_t/(1-\beta_1^t)$, $\hat{v}_t = v_t/(1-\beta_2^t)$.

\end{document}
